### 2017-07-28 LaTeX/Turnover.tex
Minor Sally edits added up until tight linkage section

\documentclass[12pt]{article}
%\documentclass{nature}

% Including pdf figures
\usepackage{graphicx}
\graphicspath{ {../Plots/} }
\usepackage{pdfpages}
%really place a figure in a location
\usepackage{float}
%Overrun caption
\usepackage[CaptionAfterwards]{fltpage}
% Math stuff
\usepackage{amsmath}
\usepackage{stix}
% Bibliographies
\usepackage[numbers]{natbib}
\bibpunct{(}{)}{,}{a}{}{;}

\usepackage[flushleft]{threeparttable}

\usepackage[font={scriptsize}]{caption}

\usepackage{lineno} %gives line numbers with \lineno command

\usepackage{setspace}
\onehalfspace

\usepackage{tikz}%for putting words on figures
\usetikzlibrary{positioning}%for relative positioning

\begin{document}

\title{Haploid Selection, Sex Ratio Bias, and Transitions Between sex-determination systems}
\author{Michael F Scott$*^1$, Matthew M Osmond$*^2$, and Sarah P Otto$^2$}
\date{}
\maketitle
\noindent
$*$ These authors contributed equally to this work

\noindent
$^1$ Department of Botany, University of British Columbia, \#3529 - 6270 University Boulevard, Vancouver, BC, Canada V6T 1Z4

\noindent
$^2$ Department of Zoology, University of British Columbia, \#4200 - 6270 University Boulevard, Vancouver, BC, Canada V6T 1Z4

\noindent
email: mfscott@biodiversity.ubc.ca, mmosmond@zoology.ubc.ca

\noindent
Contributions:

\newpage
\linenumbers
\modulolinenumbers[2]

\begin{abstract}
Sex-determination systems are remarkably dynamic; many taxa display shifts in the location of sex-determining loci or the evolution of entirely new sex-determining systems. 
Predominant theories for why we observe such transitions %in which new sex-determining systems spread by selection 
%involve sex-ratio selection or sex-specific selection (e.g., sexually antagonistic selection).
%These studies 
generally conclude that novel sex-determining systems are favoured by selection if they equalise the sex ratio or increase linkage with a sexually-antagonistic locus. 
We use population genetic models to extend these theories in two ways: 
(1) We explicitly consider how selection on very tightly sex-linked loci influences the spread of novel sex-determiners. 
We find that tightly sex-linked genetic variation can favour the spread of new sex-determination systems in which the heterogametic sex changes (XY to ZW or ZW to XY) and the new sex-determining region is less closely linked (or unlinked) to the sex-linked locus under selection; a result that is not found with loose sex-linkage. 
%With tight linkage, selection can maintain equilibria at which the allele fixed on the sex-limited chromosome (Y or W) is simultaneously favoured in the homogametic sex (XX females or ZZ males) such that heterogametic transitions (XY to ZW or ZW to XY) are favoured by selection. 
%Alternatively, a neo-W (neo-Y) locus can invade when it allows greater specialism on alleles beneficial in females (male) compared to an ancestral XX female (ZZ male), 
(2) We also consider selection upon haploid genotypes either during gametic competition (e.g., pollen/sperm competition) or meiosis (i.e., non-Mendelian segregation); selective processes that typically occur in one sex or the other. 
%As well as having sex-specific fitness consequences, haploid selection can cause the zygotic sex ratio to become biased because sex ratios are determined by the production and fertilization success of X- versus Y-bearing pollen/sperm (or Z- versus W-bearing ovules/eggs). 
%Consequently, selection for XY to ZW transitions and ZW to XY transitions can be assymetrical when linkage between the ancestral sex-determining locus and a locus under haploid selection is tight, in which case ancestral sex ratio biases can be strong.
%In addition, when linkage between the ancestral sex-determining locus and a locus under haploid selection is tight, sex ratio biases can be strong such that there is an asymmetry between invasion of ancestrally XY and ancestrally ZW systems because, e.g., haploid selection in males only causes biased zygotic sex ratios in an ancestrally XY system. 
With haploid selection, we again find that transitions between male and female heterogamety can occur even if the new sex-determining region is less closely linked to the locus under selection, and when linkage is tight haploid selection in the heterogametic sex can cause strong sex ratio bias, which may increase or decrease with the spread of new sex chromsomes. 
These results indicate that favourable associations that develop between the ancestral sex-determining locus and selected loci can be broken during the spread of a new sex-determining region. 
%Such transitions are not possible with diploid selection alone, in which case tighter linkage increases the fitness of both males and females. 
%Notably, we find that the spread of new genetic sex-determination systems is not affected by sex ratio biases that are caused by haploid selection. 
%A surprising result given that other determinants of sex allocation typically experience strong Fisherian sex-ratio selection to equalize sex ratios. 
%Overall, we demonstrate when and how the evolution of sex determining systems is impacted by sex ratio biases resulting from haploid selection and we show that new sex-determining regions can be favoured even if they are not closely linked with a locus under selection. 
Overall, our models provide new predictions for the types of selection and the genomic location of loci that can drive transitions between sex-determination systems. %and for the factors determining the phylogenetic pattern of sex-determination systems. 
\end{abstract}

%abstract word count: $\approx$ 350

\newpage

\section*{Introduction}

%CHECK THROUGHOUT: use of sex-determination SYSTEMS versus MECHANISMS. 
%CHECK THROUGHOUT: use of haploid selection, gametic selection, and meiotic drive.

Animals and angiosperms exhibit extremely diverse sex-determination systems \citep[reviewed in][]{Bull:1983vi,Charlesworth:2010it,Beukeboom:2014vb,Bachtrog:2014bx}. 
Among species with genetic sex determination of diploid sexes, some taxa have heterogametic males (XY) and homogametic females (XX), including %non-monotreme? 
mammals and most dioecious plants \citep{Ming:2011iy}; whereas other taxa have homogametic males (ZZ) and heterogametic females (ZW), including Lepidoptera and birds. 
Within several taxa, the chromosome that harbours the master sex-determining region changes. 
For example, transitions of the master sex-determining gene between chromosomes or the evolution of new master sex-determining genes have occurred in Salmonids \citep{Li:2011fm,Yano:2012di}, Diptera \citep{Vicoso:2015hf}, and \textit{Oryzias} \citep{Myosho:2012fv}.
%Presentation at evolution found neo sex chromosome in birds, doesn't seem to be published or bioRxiv yet (consider pers com.), title: A previously unknown neo-sex chromosome mediates plumage divergence and speciation in hybridizing birds Jason Sardell; Elizabeth Cooper; J. Albert C. Uy. I wonder if this is actually a fusion event though?
In addition, many gonochoric clades with genetic sex determination exhibit transitions between male (XY) and female (ZW) heterogamety, including snakes \textcolor{red}{(Gamble et al. 2017, Current Biology)}, lizards \citep{Ezaz:2009tk}, eight of 26 teleost fish families \citep{Mank:2006bt}, true fruit flies \citep[Tephritids,][]{Vicoso:2015hf}, amphibians \citep{Hillis:1990gu}, the angiosperm genus \textit{Silene} \citep{Slancarova:2013dq}, and Coleoptera and Hemiptera \citep[][plate 2]{Beukeboom:2014vb}.
Indeed, in some cases, both male and female heterogametic sex-determination systems can be found in the same species, as exhibited by some cichlid species \citep{Ser:2010iq} and \textit{Rana rugosa} \citep{Ogata:2007jm}.
In addition, multiple transitions have occurred between genetic and environmental sex-determination systems, e.g., in reptiles and fishes \citep{Conover:1987in,Mank:2006bt,Pokorna:2009ui,Ezaz:2009tk,Pen:2010kk,Holleley:2015hc}.

%NOTE: ``Bull & Charnov (1977) hypothesized that a new sex-determining gene can rapidly increase and become fixedin a population if it is linked to a gene with high adaptivevalue, and finally cause a change of the heterogametic sex.'' The Bull and Charnov paper is also the one in which the set of neutral equilibria between XY and ZW are identified (where sex ratios are equal)
%NOTE: Vuilleumier et al. 2007 also find polymorphic sex determination can be stable.

%For example, closely related species in the Medaka lineage (cite Takenhana et al. 2008) and Tilapiinae (cite Lee et al 2004, Cnaani et al 2008), citations in Beukeboom.

%\textcolor{red}{Brief description of sex ratio adjustment and sexual antagonism theories:}

Predominant theories accounting for the spread of new sex-determination systems by selection involve fitness differences between sexes (e.g., sexually antagonistic selection) or sex-ratio selection. 
%\citet{Bull:1977wt} shows that direct selection on new sex determination loci can allow them to spread and switch between male and female heterogamety. 
\citet{vanDoorn:2007eu,vanDoorn:2010hu} show that new sex-determining loci can be favoured if they arise in closer linkage with a locus that experiences sexual antagonism. 
Tighter linkage allows a stronger favourable association to build up between a male-beneficial allele, and a neo-Y chromosome, for example. 
Such associations can favour a new master sex-determining gene on a new chromosome \citep{vanDoorn:2007eu} and can also favour a transition between male and female heterogamety \citep[e.g., a ZW to XY transition,][]{vanDoorn:2010hu}. 
However, any sexually-antagonistic loci that are more closely linked to the ancestral sex-determination locus will develop similar, favourable associations and hinder the spread of a new sex-determination system. 
%Here we extend these studies by explicitly calculating the the equilibrium allele frequencies of loci that are very tightly linked to the ancestral sex-determining region.

The sex ratio is directly affected by the sex-determination system, and it has therefore been suggested that sex-ratio selection is a dominant force in the evolution of sex determination (e.g., \citealt{Bull:1983vi}, p 66-67; \citealt{Beukeboom:2014vb}, Chapter 7). 
`Fisherian' sex-ratio selection favours a 1:1 zygotic sex ratio when assuming that males and females are equally costly to produce \citep{Fisher:1930wy,Charnov:1982wg}.
This follows from the fact that, for an autosomal locus, half of the genetic material is inherited from a male and half from a female \citep{West:2009we}. 
Thus, if the population sex ratio is biased towards one sex, the average per-individual contribution of genetic material to the next generation from the opposite sex is greater. 
Therefore, a mutant that increases investment in the rarer sex will spread via the higher per-individual contributions made by that sex. 
%The default mode of sex-ratio selection is `Fisherian' sex-ratio selection, which favours equal investment in male and female offspring \citep[i.e., a 1:1 zygotic sex ratio when assuming that males and females are equally costly to produce,][]{Fisher:1930wy,Charnov:1982wg}. 
%Given that the sex-determination system can directly affect the sex ratio, we might expect Fisherian sex-ratio selection to influence the spread of new sex-determination systems. 
In the case of sex-chromosome evolution, \citet{Kozielska:2010vm} consider systems in which the ancestral sex chromosomes experience meiotic drive (e.g., where driving X or Y chromosomes are inherited disproportionately often), which causes sex ratios to become biased \citep{Hamilton:1967ts}. 
They find that new, unlinked sex-determining loci (masculinizing or feminizing mutations, i.e., neo-Y or neo-W loci) can then spread, which restore an even sex ratio.

%\textcolor{red}{We add haploid selection:}
Here we extend current theory %on turnover between sex-determining systems 
by using mathematical models to find the conditions under which new sex-determination systems spread when individuals experience selection at both diploid and haploid stages. 
Even in animal and plant species that have much larger and more conspicuous diploid phases than haploid phases, many loci experience significant haploid selection through gamete competition and/or meiotic drive \citep{Mulcahy:1996ha,JOSEPH:2004haa}.
We use the term `meiotic drive' to refer to the biased (non-Mendelian) segregation of genotypes during gamete production (from one parent) and the term `gametic competition' to refer to selection upon haploid genotypes within a gamete/gametophyte pool (potentially from multiple parents); the term `haploid selection' encompasses both processes. 
Meiotic drive generally occurs either during the production of male or female gametes only \citep{Ubeda:2005gw,Lindholm:2016cw}.
Because there are typically many more pollen/sperm than required for fertilization, gametic competition is also typically sex specific, occurring primarily among male gametes.
Gametic competition may be particularly common in plants, in which 60-70\% of all genes are expressed in the male gametophyte and these genes exhibit stronger signatures of selection than random genes \citep{Borg:2009jpa,Arunkumar:2013iq,Gossmann:2014dua}.
In addition, artificial selection pressures applied to male gametophytes are known to cause a response to selection \citep[e.g.,][]{Hormaza:1996cv,Ravikumar:2003uo,Hedhly:2004iv,Clarke:2004ir}. 
A smaller proportion of genes are thought to be expressed and selected during competition in animal sperm, although precise estimates are uncertain \citep{Zheng:2001fi,JOSEPH:2004haa,Vibranovski:2010et}. 
Recent studies have demonstrated that sperm competition can alter haploid allele frequencies and increase offspring fitness \citep{Immler:2014im} \textcolor{red}{(Alavioon et al. 2017). 
Genetic mapping experiments, which are typically designed to minimize selection in diploids, have revealed segregation distortion in various species, including mice, Drosophila, Rice, Maize, Wheat, Barley, Cotton... 
In some of these cases, biased segregation has been shown to be attributable to meiotic drive and/or gametic selection \textcolor{red}{(Leppala et al. 2013, Didion et al. 2015, 2016 Xu et al 2013 (rice), Fishman...)}. }
% from Bai et al. (2016) ``[Segregation distortion] phenomenon is commonly detected via genetic mapping and has been documented in various species, including mouse (Eversley et al., 2010; Casellas et al., 2012), Drosophila (Phadnis and Orr, 2009; Larracuente and Presgraves, 2012; McDermott and Noor, 2012), Tigriopus (Pritchard et al., 2011), rice (Koide et al., 2012; Reflinur et al., 2014; Xu et al., 2014), maize (Tang et al., 2013), and cotton (Yu et al., 2011; Hulse-Kemp et al., 2015).'' Also more in that paper that might be useful. 
%Drosophila http://www.biorxiv.org/content/early/2017/02/06/106336

There are various ways in which a period of haploid selection could influence transitions between sex-determination systems. 
If we assume that haploid selection at any particular locus predominantly occurs in one sex (e.g., meiotic drive during spermatogenesis), then such loci experience a form of sex-specific selection. 
In this respect, we might expect that haploid selection would affect transitions between sex-determination systems in a similar manner to sex-specific diploid selection \citep[as explored by][]{vanDoorn:2007eu,vanDoorn:2010hu}. 
That is, new masculinizing mutations (neo-Y chromosomes) could be favoured via associations with alleles that are beneficial in the male haploid stage. 
On the other hand, sex ratios can also become biased by linkage between the sex-determining region and a locus that harbours genetic variation in haploid fitness. 
For example, there are several known cases of sex-ratio bias caused by sex-linked meiotic drive alleles \citep[][Chapter 3]{Burt:2006} or selection among X- and Y-bearing pollen \citep{Lloyd:1974tz,Conn:1981uw,Stehlik:2005ul,Stehlik:2006to,Field:2012fd,Field:2013cc}. 
It is not immediately clear how the spread of new sex-determination systems would be influenced by the combination of sex-ratio biases and associations between haploid selected loci and sex-determining regions.

Our models have two important new features. 
Firstly, when considering loci that are under selection and also in very tight linkage with the ancestral sex-determining region we explicitly calculate equilibrium allele frequencies.
This allows us to show that transitions between male and female heterogamety can evolve even when the neo-sex-determining locus is less closely linked to a locus under selection and therefore disrupts favourable ancestral associations between sex and the alleles selected in that sex.
%Such transitions are not favoured in models lacking tight linkage and/or haploid selection. 
Secondly, we allow sex-specific haploid selection to occur on a locus in tight or loose linkage with the ancestral sex-determining region. 
We find that sex-ratio biases caused by haploid selection can exert Fisherian sex-ratio selection upon novel sex-determiners but that their spread is also determined by selection on genetically-associated alleles. 
%Indeed, it is only when haploid-selected loci are tightly linked to the ancestral sex-determining region (and so sex-ratio biases are initially large) that we see an asymmetry between selection for XY to ZW transitions and ZW to XY transitions (e.g., because haploid selection in males only causes biased zygotic sex ratios in an ancestrally XY system).
Consequently, it is possible for selection on linked alleles to drive turnover between sex-determining systems despite causing transitory or even permanent increases in sex-ratio bias. 
%In addition, when linkage between the ancestral sex-determining locus and a locus under haploid selection is tight, sex ratio biases can be strong such that there is an asymmetry between invasion of ancestrally XY and ancestrally ZW systems because, e.g., haploid selection in males only causes biased zygotic sex ratios in an ancestrally XY system.

%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Model}
%%%%%%%%%%%%%%%%%%%%%%%%%

\textcolor{red}{Change all  $\alpha^\Hermaphrodite$ to $(1+\alpha_{\Delta}^\Hermaphrodite)$.}

%genetic system
We consider transitions between ancestral and novel sex-determining systems using a three-locus model, each locus having two alleles. 
Locus \textbf{X} is the ancestral sex-determining region, with alleles $X$ and $Y$ (or $Z$ and $W$).
Locus \textbf{A} is a locus under selection, with alleles $A$ and $a$.
Locus \textbf{M} is a novel sex-determining region, at which the null allele ($M$) is initially fixed in the population such that sex of zygotes is determined by the genotype at the ancestral sex-determining region, \textbf{X}; $XX$ genotypes become females and $XY$ become males (or $ZW$ become females and $ZZ$ become males). 
To evaluate the evolution of new sex-determination systems, we consider the invasion, fixation, maintenance, and/or loss of novel sex-determining alleles ($m$) at the \textbf{M} locus. 
%We only do invasion analytically though (is this sentence still ok?)
We assume that the \textbf{M} locus is epistatically dominant over the \textbf{X} locus such that zygotes with at least one $m$ allele develop as females with probability $k$ and as males with probability $1-k$, regardless of the \textbf{X} locus genotype.
With $k=0$, the $m$ allele is a masculinizer (i.e., a neo-Y) and with $k=1$ the $m$ allele is a feminizer (i.e., a neo-W).
With intermediate $k$, we can interpret $m$ as an environmental sex determination (ESD) allele, such that zygotes develop as females in a proportion ($k$) of the environments they experience. 
We also analyze a model of maternally-controlled environmental sex-determination, where mothers with at least one $m$ allele produce daughters with probability $k$.

%life-cycle
In each generation, we census the genotype frequencies in male and female gametes/gametophytes (hereafter gametes) before gametic competition. 
A full description of our model, including recursion equations, is given in the Appendix. 
First, competition occurs among male gametes (sperm/pollen competition) and among female gametes (egg/ovule competition) separately. 
Selection during gametic competition depends on the \textbf{A} locus genotype, relative fitnesses are given by $w_A^\Hermaphrodite$ and $w_a^\Hermaphrodite$ ($\Hermaphrodite \in \{\female,\male\}$; see table \ref{tab:fitnesstable}). %$w_{A}^\male$ and $w_{a}^\male$ for male gametes and $w_{A}^\female$ and $w_{a}^\female$ for female gametes, see table \ref{tab:fitnesstable}. 
We assume that all gametes compete for fertilization during gametic competition, which assumes a polygamous mating system. 
Gametic competition in monogamous mating systems is, however, equivalent to meiotic drive in our model (described below), as both only alter the frequency of gametes produced by heterozygotes. 
After gametic competition, random mating occurs between male and female gametes.
The resulting zygotes develop as males or females, depending on their genotypes at the \textbf{X} and \textbf{M} loci. %(and the \textbf{M} genotype of their mother in the case of maternal control).
Diploid males and females then experience selection, with relative fitnesses $w_{AA}^{\Hermaphrodite}$, $w_{Aa}^{\Hermaphrodite}$, and $w_{aa}^{\Hermaphrodite}$. %where $g$ is the diploid genotype at the \textbf{A} locus ($g \in \{AA, Aa, aa\}$).  
%Diploids compete with others of the same sex, where selection depends on the sex of the individual and its genotype at the \textbf{A} locus.
The next generation of gametes is produced by meiosis, during which recombination and sex-specific meiotic drive can occur. 
Recombination (i.e., an odd number of cross-overs) occurs between loci \textbf{X} and \textbf{A} with probability $r$, between loci \textbf{A} and \textbf{M} with probability $R$, and between loci \textbf{X} and \textbf{M} with probability $\chi$.
Any linear order of the loci can be modelled with appropriate choices of $r$, $R$, and $\chi$ (see Table \ref{tab:chisubstitutions}). 
Individuals that are heterozygous at the \textbf{A} locus may experience meiotic drive; a gamete produced by $Aa$ heterozgotes of sex $\Hermaphrodite$ bear allele $A$ with probability $\alpha^\Hermaphrodite$. 
Thus, the \textbf{A} locus can experience sex-specific gametic competition, diploid selection, and/or meiotic drive.

\begin{table}[ht]
\centering
\smallskip
\caption{Relative fitness of different genotypes in sex $\Hermaphrodite \in \{\female,\male\}$}
\begin{tabular}{l c }
\hline\hline
  Genotype & Relative fitness during gametic competition \\ [0.5ex] \hline
  A & $w_{A}^\Hermaphrodite = 1+t^\Hermaphrodite$ \\
  a & $w_{a}^\Hermaphrodite = 1$ \\ [0.5ex] \hline
  Genotype & Relative fitness during diploid selection \\ [0.5ex] \hline
  AA & $w_{AA}^\Hermaphrodite = 1+ s^\Hermaphrodite$ \\
  Aa & $w_{Aa}^\Hermaphrodite = 1+h^\Hermaphrodite s^\Hermaphrodite$ \\
  aa & $w_{aa}^\Hermaphrodite = 1$ \\ [0.5ex] \hline
  Genotype & Transmission during meiosis in $Aa$ heterozygotes \\ [0.5ex] \hline
  A & $\alpha^\Hermaphrodite=1/2+\alpha_{\Delta}^{\Hermaphrodite}/2$ \\
  a & $1-\alpha^\Hermaphrodite=1/2-\alpha_{\Delta}^{\Hermaphrodite}/2$ \\
  \hline \hline
  \label{tab:fitnesstable}
 \end{tabular}
\end{table}

%Info moved above
%what were going to do, briefly
%With allele $M$ fixed at the \textbf{M} locus, sex is determined by locus \textbf{X} and an equilibrium is reached at locus \textbf{A}. %where, importantly, locus \textbf{A} can be stably polymorphic. %the dynamics can be described by the frequency of $A$ in gametes/gametophytes from the homogametic sex and the frequencies of $A$ in the two types of gametes/gametophytes from the heterogametic sex.
%We then examine the ability of a rare, novel sex-determiner, $m$, to increase in frequency from this equilibrium.
%Numerical simulations are used to examine when an invading mutation goes to fixation.

%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
%%%%%%%%%%%%%%%%%%%%%%%%%

The model outlined above describes both ancestrally-$XY$ and ancestrally-$ZW$ sex-determination systems if we relabel the two sexes as being ancestrally `heterogametic' or ancestrally `homogametic'. 
Without loss of generality, we primarily refer to the ancestrally heterogametic sex as male and the ancestrally homogametic sex as female.
That is, we describe an ancestral XY sex-determination system but our model is equally applicable to an ancestral ZW sex-determination system (relabelling the ancestrally-heterogametic sex as female and the ancestrally-homogametic sex as male).

%\subsection*{Turnover between sex-determination systems}

\subsubsection*{Generic invasion by a neo-Y or neo-W}

The evolution of a new sex-determination system requires that a rare mutant allele at the novel sex-determining locus, $m$, increases in frequency when rare. 
The spread of a rare mutant $m$ at the \textbf{M} locus is determined by the leading eigenvalue, $\lambda$, of the system of eight equations describing the frequency of eggs and sperm carrying the $m$ allele in the next generation (equations \ref{eq:recursions}). %, evaluated at the above equilibrium.
%Thus, the system describing the change in frequency of the $m$ allele consists of eight recursion equations. 
This system simplifies substantially in a number of cases of interest. 
Dominant neo-Y (when $k=0$) or neo-W alleles (when $k=1$) are only found in male diploids (neo-Y) or female diploids (neo-W) such that their growth rate ultimately depends only on the change in frequency of $m$-bearing gametes produced by males or by females, respectively. 
Furthermore, if the $m$ allele is fully epistatically dominant over the ancestral sex-determining system, phenotypes are not affected by the genotype at the ancestral sex-determining region (\textbf{X} locus). 
Thus, the invasion of rare dominant neo-Y or neo-W alleles is determined by the largest eigenvalue that solves a quadratic characteristic polynomial,
$\lambda^2+ b \lambda + c =0$.
%\begin{equation}
%\lambda^2+ b \lambda + c =0
%\label{eq:charpoly_neoY}
%\end{equation}
%
%\noindent
%where $b$ is the average of the growth rates of the two haplotypes that carry the $m$ allele ($mA$ and $ma$), $b=(\lambda_{mA}+\lambda_{ma})/2$, and $c$ also involves the fitness of $m$ alleles when they recombine onto the other \textbf{A} background in a heterozygote, $c=\lambda_{mA}\lambda_{ma}+\rho_{mA}\rho_{ma}$ (see table \ref{tab:haplotype_growth}).
%Here, $b= - (\lambda_{mA} + \lambda_{ma})$ and $c = \lambda_{mA} \lambda_{ma} -\rho_{mA} \rho_{ma}$, where $\lambda_{mi}$ is the (multiplicative) growth rate of mutant haplotypes on background $i\in\{A,a\}$, accounting for loss due to recombination, and $\rho_{mi}$ is the rate of addition of mutant haplotypes onto background $i\in\{A,a\}$ due to recombination (see table \ref{tab:haplotype_growth}).
Here, $b= - (\lambda_{mA} + \lambda_{ma})+(\rho_{mA} + \rho_{ma})$ and $c = (\lambda_{mA}-\rho_{mA}) (\lambda_{ma}-\rho_{ma}) -\rho_{mA} \rho_{ma}$, where $\lambda_{mi}$ is the multiplicative growth rate of mutant haplotypes on background $i\in\{A,a\}$, without accounting for loss due to recombination, and $\rho_{mi}$ is the rate at which mutant haplotypes on background $i\in\{A,a\}$ recombine onto the other \textbf{A} locus background in heterozygotes (see Table \ref{tab:haplotype_growth}).
The $\lambda_{mi}$ and $\rho_{mi}$, and thus the spread of the mutant $m$ allele, depend on the frequency of alleles at the \textbf{A} and \textbf{X} loci in the ancestral population. 
In the ancestral population, it is convenient to follow the frequency of the $A$ allele among female gametes (eggs), $p^\female_X$, and among X-bearing, $p^\male_X$, and among Y-bearing, $p^\male_Y$, male gametes (sperm/pollen). 
We also track the fraction of male gametes that are Y-bearing, $q$, which may deviate from $1/2$ due to meiotic drive in males.

%\textcolor{red}{I now remove the zygotic sex ratio $\zeta$ from the mean fitnesses. The mean fitnesses will have to be adjusted in a corresponding way.}
%
%\textcolor{red}{CHECK: DO THE HAPLOID MEAN FITNESSES HAVE TO BE IN THE DENOMINATOR? i think so, based on definitions in table S.2 (mmo)}

% as there are now two 2s in the lambdas of table 2. perhaps the 2 in the denominator is needed because the frequency of males sums to one as does the frequency of females?}

\begin{threeparttable}[ht]
\centering
\smallskip
%\caption{Parameters determining invasion (equation \ref{eq:charpoly_neoY}) for neo-Y or neo-W alleles}
\caption{Parameters determining invasion of mutant neo-Y and neo-W alleles into an ancestrally XY system}
\begin{tabular}{l}
\hline\hline
   \noalign{\vskip 0.5ex}
   neo-Y ($k=0$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\lambda_{mA} = \left\{ 2 (1- \zeta) \right\}^{-1} \left[p_{X}^{\female} w_{A}^{\female} w_{A}^{\male} w_{AA}^{\male} + 2(1-p_{X}^{\female}) w_{a}^{\female} w_{A}^{\male} w_{Aa}^{\male} \alpha^\male \right]/ \left( \bar{w}_H^\female \bar{w}_H^\male \bar{w}^\male \right) $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\lambda_{ma} = \left\{ 2 (1- \zeta) \right\}^{-1} \left[(1-p_{X}^{\female}) w_{a}^{\female} w_{a}^{\male} w_{aa}^{\male} + 2p_{X}^{\female} w_{A}^{\female} w_{a}^{\male} w_{Aa}^{\male}(1 - \alpha^\male) \right]/ \left( \bar{w}_H^\female \bar{w}_H^\male \bar{w}^\male \right) $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{mA} = R \left\{ 2 (1- \zeta) \right\}^{-1} \left[ 2(1-p_{X}^{\female}) w_{a}^{\female} w_{A}^{\male} w_{Aa}^{\male} \alpha^{\male} \right]/  \left( \bar{w}_H^\female \bar{w}_H^\male \bar{w}^\male \right)   $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{ma} = R \left\{ 2 (1- \zeta) \right\}^{-1} \left[  2 p_{X}^{\female} w_{A}^{\female} w_{a}^{\male} w_{Aa}^{\male} (1 - \alpha^{\male}) \right]/ \left( \bar{w}_H^\female \bar{w}_H^\male \bar{w}^\male \right)  $\\ [1ex] \hline 
  \noalign{\vskip 0.5ex}
  neo-W ($k=1$) \\ [0.5ex] \hline \noalign{\vskip 1ex}
  $\lambda_{mA} = \left( 2 \zeta \right)^{-1} \left[ \bar{p}^{\male} w_{A}^{\male} w_{A}^{\female} w_{AA}^{\female}+2(1-\bar{p}^{\male}) w_{a}^{\male} w_{A}^{\female} w_{Aa}^{\female}\alpha^\female\right]/ \left(\bar{w}_H^\female \bar{w}_H^\male \bar{w}^\female \right) $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\lambda_{ma} = \left( 2 \zeta \right)^{-1} \left[ (1-\bar{p}^{\male}) w_{a}^{\male} w_{a}^{\female} w_{aa}^{\female}+2\bar{p}^{\male} w_{A}^{\male} w_{a}^{\female} w_{Aa}^{\female}(1-\alpha^\female)\right] / \left(\bar{w}_H^\female \bar{w}_H^\male \bar{w}^\female \right) $ \\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{mA} = R \left( 2 \zeta \right)^{-1} \left[ 2(1-\bar{p}^{\male}) w_{a}^{\male} w_{A}^{\female} w_{Aa}^{\female} \alpha^{\female} \right] / \left(\bar{w}_H^\female \bar{w}_H^\male \bar{w}^\female \right)  $\\ [0.5ex] \noalign{\vskip 0.5ex}
  $\rho_{ma} = R \left( 2 \zeta \right)^{-1} \left[ 2\bar{p}^{\male} w_{A}^{\male} w_{a}^{\female} w_{Aa}^{\female} (1-\alpha^{\female}) \right] / \left(\bar{w}_H^\female \bar{w}_H^\male \bar{w}^\female \right) $\\ [1ex] 
  \hline \hline 
   \end{tabular}
      \begin{tablenotes}
      \scriptsize
      \item $\bar{p}^{\male}=(1-q)p_{X}^{\male} + qp_{Y}^{\male}$ is the average frequency of the $A$ allele among X- and Y-bearing male gametes.
%      \item $R$ is the probability of recombination between loci \textbf{A} and \textbf{M}.
      \item $\zeta$ is the zygotic sex ratio (fraction female)
      \item $\bar{w}^\Hermaphrodite$ is the mean fitness of diploids of sex $\Hermaphrodite$, see Table \ref{tab:meanfitnesses}
      \item $\bar{w}_H^\Hermaphrodite$ is the mean fitness of haploids from sex $\Hermaphrodite$, see Table \ref{tab:meanfitnesses}
%      \item $\bar{w}_H^\female = p_X^\female w_A^\female + (1-p_X^\female) w_a^\female$ is the mean fitness of female gametes during gametic competition
%      \item $\bar{w}_H^\male = \bar{p}^{\male} w_A^\male + (1-\bar{p}^{\male}) w_a^\male$ is the mean fitness of male gametes during gametic competition
%       \item $\bar{w}^\female = \{ p_X^\female w_A^\female (1-q) p_X^\male w_A^\male w_{AA}^\female + (1 - p_X^\female) w_a^\female (1-q) p_X^\male w_A^\male w_{Aa}^\female + p_X^\female w_A^\female (1-q) (1 - p_X^\male) w_a^\male w_{Aa}^\female + (1-p_X^\female) w_a^\female (1-q) (1 - p_X^\male) w_a^\male w_{aa}^\female \} / \{ \bar{w}_H^\female \bar{w}_H^\male\}$ is the mean fitness of females during diploid competition
%       \item $\bar{w}^\male = \{ p_X^\female w_A^\female q p_Y^\male w_A^\male w_{AA}^\male + (1 - p_X^\female) w_a^\female q p_Y^\male w_A^\male w_{Aa}^\male + p_X^\female w_A^\female q (1 - p_Y^\male) w_a^\male w_{Aa}^\male + (1-p_X^\female) w_a^\female q (1 - p_Y^\male) w_a^\male w_{aa}^\male \} / \{ \bar{w}_H^\male \bar{w}_H^\male\}$ is the mean fitness of males during diploid competition
    \end{tablenotes}
  \label{tab:haplotype_growth}
\end{threeparttable}
\\

%We are particularly concerned with the conditions under which a rare neo-sex-determining allele increases in frequency, which occurs when the largest eigenvalue, $\lambda$, is greater than one. 
%If one or both of $\lambda_{mA}$ and $\lambda_{ma}$ are greater than one, indicating that at least one haplotype increases in frequency faster than recombination breaks it apart, invasion will always occur ($\lambda>1$). 
%Otherwise, invasion requires
%\begin{equation}
%(\lambda_{mA}-1)(\lambda_{ma}-1) - \rho_{mA} \rho_{ma}<0.
%\end{equation}

We are particularly concerned with the conditions under which a rare neo-sex-determining allele increases in frequency, which occurs when the largest eigenvalue, $\lambda$, is greater than one. 
If the average change in frequency of the two haplotypes that carry the $m$ allele ($Am$ and $am$) is positive, invasion will always occur, i.e., if $(\lambda_{mA} + \lambda_{ma})/2 > 1$ then $\lambda > 1$. 
If neither haplotype increases in frequency ($\lambda_{mA}, \lambda_{ma} < 1$), the $m$ allele will not invade. 
Otherwise, the new sex-determining allele increases in frequency on one \textbf{A} background and declines on the other, and invasion requires 
\begin{equation}\label{eq:lambdasGen}
%\frac{\rho_{mA}}{\lambda_{mA}-1}+\frac{\rho_{ma}}{\lambda_{ma}-1}>0.
\rho_{ma}\left(\lambda_{mA}-1\right)+\rho_{mA}\left(\lambda_{ma}-1\right)>0.
\end{equation}
\noindent For example, if we assume that only the $mA$ haplotype has a positive growth rate ($\lambda_{ma}<1<\lambda_{mA}$), the second term on the left-hand side of \eqref{eq:lambdasGen} is negative and invasion requires that the growth rate of $mA$ haplotypes and the rate at which they are produced by recombination is sufficiently large relative to that of $ma$ haplotypes.
In other words, invasion requires that the average growth rate of the two haplotypes, weighted by the rates they are created by recombination, is positive.
%the number of recombinants from the fitter $mA$ haplotypes is larger than that of the less fit $ma$ haplotypes, relative to their growth rates.

Table \ref{tab:haplotype_growth} illustrates a number of key points about the invasion of neo-Y and neo-W mutations. 
First, Fisherian sex-ratio selection will favour the spread of a neo-Y if the ancestral zygotic sex ratio is biased towards females, $\zeta>1/2$ (i.e., $\zeta>1/2$ causes the first factor of the $\lambda_{mi}$ to be greater than one for a neo-Y and vice versa for a neo-W).
However, the spread of a neo-Y (neo-W) also depends on the male (female) fitness of associated alleles (terms in square brackets). %, see terms in curly brackets. 
Second, invasion by a neo-Y (neo-W) allele does not directly depend on the fitness of female (male) diploids (for a given set of equilibrium allele frequencies).
%involve any female (male) diploid selection terms ($w_{g}^{\female}$)
This is because a dominant neo-Y (neo-W) is always found in males (females), and therefore the frequency of the neo-Y (neo-W) allele, $m$, only changes in males (females). 
%Therefore, invasion by a neo-Y allele does not involve any female diploid selection terms ($w_{g}^{\female}$). 
%Similarly, invasion by a neo-W is driven by the fitness of female gametes and diploids and does not involve any direct selection in male diploids. 
Finally, invasions by a neo-Y and a neo-W are qualitatively different.
This is because a gamete with the ancestral- or neo-Y always pairs with a female gamete containing an X, and both develop into males.
By contrast, a gamete with a neo-W can pair with an X or Y male gamete, developing into a female, while female gametes without the neo-W can become female (when paired with X) or male (when paired with Y).
Consequently, the types of females produced differ in the frequency of $A$ alleles they obtain from mating.
%\textcolor{blue}{nice}
%the diploid fitness terms in Table \ref{tab:haplotype_growth} are weighted by the probability of producing those genotypes through matings with gametes of the opposite sex. 
%For example, matings between a neo-Y-bearing male gamete and an $A$-bearing female gamete occur with probability $p_{X}^{\female} w_{A}^{\female}/\bar{w}_{H}^{\female}$. 
%This is the same probability that an ancestral-Y-bearing male gamete will inherit an $A$ allele via mating with a female gamete. 
%On the other hand, 
%The probability that a neo-W bearing female gamete mates with an $A$-bearing male gamete is $\bar{p}^{\male} w_{A}^{\male}/\bar{w}_{H}^{\male}$, where $\bar{p}^{\male}=p_{Y}^{\male}q+p_{X}^{\male}(1-q)$ is the frequency of the A allele among both X- and Y-bearing male gametes. 
%That is, in the case of a neo-W, female diploids can result from matings with either an X-bearing or a Y-bearing sperm, resulting in zygotes that will develop as females. 
%However, females that do not carry the neo-W only result from matings with X-bearing sperm. 
%Therefore, eggs with and without a neo-W can differ in the frequency of $A$ alleles they obtain from matings with male gametes. 
%Invasions by a neo-Y and a neo-W differ in this respect because sperm with or without a neo-Y allele both mate with X-bearing female gametes only.

% invasion is driven by the fitness of male gametes and diploids, where the latter is weighted by the chance that a female egg will give rise to that diploid. 
%For example, matings with $A$-bearing female gametes occur with probability $p_{X}^{\female} w_{A}^{\female}/\bar{w}_{H}^{\female}$. 
%Since a neo-Y is always found in males, the allele frequencies at the neo-Y (\textbf{M}) locus only change in males. 
%Therefore, invasion by a neo-Y allele does not involve any female diploid selection terms ($w_{g}^{\female}$). 
%Similarly, invasion by a neo-W is driven by the fitness of female gametes and diploids and does not involve any direct selection in male diploids.
%However, in the case of a neo-W, female diploids can result from matings with either an X-bearing or a Y-bearing sperm. 
%In either case, the zygote will develop as a female due to the presence of a neo-W. 
%For example, neo-W females will inherit an $A$ allele from a male gamete with probability $\bar{p}^{\male} w_{A}^{\male}/\bar{w}_{H}^{\male}$, where $\bar{p}^{\male}=p_{Y}^{\male}q+p_{X}^{\male}(1-q)$ is the frequency of the A allele among both X- and Y-bearing male gametes. 
%By contrast, females that do not carry the neo-W only result from matings with X-bearing sperm. 
%They will therefore inherit an $A$ from a male gamete with probability $(1-q)\bar{p}_{X}^{\male} w_{A}^{\male}/\bar{w}_{H}^{\male}$. 
%If the \textbf{A} locus is initially linked to the ancestral sex-determining locus, \textbf{X}, the frequency of the $A$ allele among X- and Y-bearing sperm can differ (equation \ref{eq:freq_diffs}). 
%Thus, eggs with and without a neo-W differ in the frequency of $A$ alleles they obtain from mating with male gametes.

In order to explicitly determine the conditions under which a rare neo-sex-determining allele spreads, we must calculate the equilibrium frequency of the $A$ allele in the ancestral population (i.e., $\hat{p}^\female_X$, $\hat{p}^\male_X$, and $\hat{p}^\male_Y$). 
Since only the \textbf{A} locus experiences selection directly, any deterministic evolution requires that there is a polymorphism at the \textbf{A} locus. 
Polymorphisms can be maintained by mutation-selection balance or transiently present during the spread of beneficial alleles. However, polymorphisms maintained by selection can maintain alleles at higher allele frequencies for longer periods. 
Here, we focus of polymorphisms maintained by selection, where the $A$ allele reaches a stable intermediate equilibrium frequency under the ancestral sex-determination system before the neo-sex-determining allele ($m$) arises. 
We can analytically calculate the allele frequency of the $A$ allele using two alternative simplifying assumptions: 
(1) the \textbf{A} locus is within (or tightly linked to) the non-recombining region around the ancestral SDR ($r \approx 0$) or (2) selection is weak relative to recombination ($s^\Hermaphrodite$, $t^\Hermaphrodite$, $\alpha_{\Delta}^\Hermaphrodite$ of order $\epsilon<<1$). 
%These equilibrium allele frequencies and their stability conditions are given in the appendix. 
%Considering diploid selection only, \citet{vanDoorn:2007eu,vanDoorn:2010hu} analytically calculated these allele frequencies assuming weak selection and relatively loose linkage (assumption 2 above). 
%We first consider the spread of neo-sex determining loci when the ancestral sex-determining locus is tightly linked to the selected \textbf{A} locus (assumption 1) with or without haploid selection. 
%We then use the weak selection approximation that was also used in \citet{vanDoorn:2007eu,vanDoorn:2010hu} except that we allow haploid selection on the \textbf{A} locus.

\textcolor{red}{Change to $\hat{p}$ throughout as we assume that allele frequencies change slowly such that lambda is unaffected}

\subsubsection*{Tight linkage with the ancestral sex-determining region}

The equilibrium allele frequencies and their stability conditions are given in the appendix, when there is complete linkage between the ancestral sex-determining region and the \textbf{A} locus ($r=0$). Here, either the $A$ allele or the $a$ allele must be fixed on the Y. 
% because selection on the Y is effectively haploid an incapably of maintaining a polymorphism in the absence of frequency dependent selection. 
Because the labelling of alleles is arbitrary, we will assume that the $a$ locus is fixed on the Y ($p^\male_Y=0$), without loss of generality . 
If there are two alleles maintained at the \textbf{A} locus, the X can either be fixed for the $A$ allele ($p^\female_X=p^\male_X=1$) or polymorphic (0<$p^\female_X, p^\male_X$<1). 
%However, because the X is present in both males and females, there are distinct selection pressures on X's in females and X's in males, which will always be paired with an $A$ allele from the Y.

A neo-Y will never invade an ancestral XY system that already has tight linkage with the locus under selection ($r=0$, for details see supplementary \textit{Mathematica} file). 
%When then neo-Y is also tightly linked ($R=0$) a neo-Y will either remain linked to the $A$ allele or to the $a$ allele and so invasion is given directly by the larger of $\lambda_{mA}$ or $\lambda_{ma}$.
A neo-Y haplotype with the same allele as the ancestral Y is neutral ($\lambda_{ma}=1$) and does not change in frequency.
The other neo-Y haplotype will not spread ($\lambda_{mA}<1$) given that the initial equilibrium is stable. 
Therefore, a neo-Y mutation cannot spread ($\lambda \leq 1$) in an ancestral XY system that is at equilibrium with all selected loci within the non-recombining region around the SDR.
In essence, through tight linkage with the \textbf{A} locus, the ancestral Y becomes strongly specialized on the allele that has the highest fitness across male haploid and diploid phases. 
Given that the ancestral Y is at this equilibrium, it is not possible for a neo-Y to create males that have higher fitness than the ancestral Y.

%\textcolor{blue}{I might be tempted to focus on the diploid selection only case here because this is a new/surprising result and we have less cool stuff to say about haploid selection beyond describing it affecting stuff (right?). I then might also leave the complication of whether there is a polymorphism on the X or not to the appendix??? Attempt 1:}

\textcolor{red}{Sally edits only looked at up to this point. Next task: figures to match with this tight linkage section.}

%\textcolor{blue}{
%Briefly, the neo-W invades under the ancestral XY sex-determining system, the X chromosome may not be able to as specialised on a female-beneficial allele
%The full characteristic polynomials are given in the appendix (equations \ref{Binvasion} and \ref{Ainvasion}).
Neo-W alleles, on the other hand, can invade an ancestral XY system under some conditions (the full invasion conditions are given in the appendix; equations \ref{Binvasion} and \ref{Ainvasion}). 
Significantly, we note that it is possible for both neo-W haplotypes to spread ($\lambda_{mA}>1$ and $\lambda_{ma}>1$), in which case neo-W invasion can occur regardless of linkage to the selected locus.
That is, selection on loci within the non-recombining region of the SDR can favour the invasion of a less closely linked neo-W (e.g., on an autosome). 
Although haploid selection can favour neo-W alleles because the ancestral sex ratio becomes male biased, this is not the only circumstance in which less tightly linked neo-W alleles invade. 
For example, unlinked neo-W alleles can invade in the absence of any haploid selection. 
This result is unexpected given the results of \cite{vanDoorn:2010hu}, who did not explicitly calculate equilibrium allele frequencies under tight linkage and generally concluded that heterogametic transitions occur when neo-sex-determining alleles are in tighter linkage with loci under sex-specific diploid selection. 
Therefore, we focus on cases where there is no haploid selection and discuss the effects of haploid selection in the appendix. 
%}

%\textcolor{blue}{
If we categorise the $a$ allele as being ancestrally `male-beneficial' via the fact that it is fixed on the Y, then $\lambda_{mA}>1$ indicates that the neo-W spreads when found with the ancestrally `female-beneficial' allele. 
Intuitively, this is possible because the ancestral X chromosome is not able to specialise on the `female-beneficial' allele due to the fact that X's are sometimes found in males. 
For example, a polymorphism of $A$ and $a$ alleles can be maintained on the X despite directional selection in favour of the $A$ allele in females ($s^\female>0$, $0<h^\female<1$) because the $a$ allele is favoured in males. 
W-$A$ haplotypes will only create females with high fitness $AA$ or $Aa$ genotypes and can therefore have higher fitness than ancestral females, which sometimes also produce $aa$ females. 
Thus, the neo-W can spread by allowing increased specialization on female beneficial alleles.
%As another example, $AA$ individuals can have low fitness in females and yet the $A$ can be fixed on the X due to strong overdominance in males favouring X$A$Y$a$ males. 
%In this case, W$A$ haplotypes produce fewer unfit $AA$ females because they produce some females through matings with Y$a$-bearing sperm. 
%}

%\textcolor{blue}{
Given that the $a$ allele can be considered ancestrally `male-beneficial' because it is fixed on the Y, it might be surprising that neo-W-$a$ haplotypes can be favoured by selection in females ($\lambda_{ma}>1$). 
Again, this occurs because ancestral X's also experience selection in males, in which they will always be paired with a Y-$a$. 
Hence, if there is overdominance in males, X-$A$ Y-$a$ males have high fitness and the $A$ allele is favoured by selection on the X in males. 
Therefore, the X can be polymorphic or even fixed for the $A$ allele despite favouring the $a$ allele during selection in females \citep{Lloyd1977,Otto2014}. 
In such cases, neo-W-$a$ haplotypes, which are never found in males, can spread because they both create more $Aa$ and $aa$ females when pairing with an X from males and they bring Y-$a$ haplotypes into females, in which case females are always $aa$. 
Indeed, it is possible for both W-$A$ and W-$a$ haplotypes to spread, as is the case when $AA$ individuals have low fitness in females yet the $A$ is fixed on the X due to strong overdominance in males.
Both neo-W-$A$ and neo-W-$a$ haplotypes then produce fewer unfit $AA$ females.
This is true for the neo-W-$A$ haplotype because it can pair with a $Y-a$ haplotype and still be female. 
%}

%\textcolor{blue}{
In Figure \ref{fig:regionplot_2a}A we show the region of parameter space within which both neo-W haplotypes invade ($\lambda_{mA}>1$ and $\lambda_{ma}>1$) when there is overdominance in females and no haploid selection \citep[corresponding to Figure 2a in][]{Otto2014}. 
Wherever both haplotypes have positive growth rates (gray region of Figure \ref{fig:regionplot_2a}), invasion by a neo-W is expected regardless of its linkage with the selected locus (i.e., even unlinked neo-W alleles can invade). 
In regions where only one haplotype can spread (white region of Figure \ref{fig:regionplot_2a}), a neo-W can invade as long as equation \eqref{eq:lambdasGen} is satisfied, which can require that the recombination rate, $R$, is small enough and yet still indicates that more loosely linked sex-determining regions can spread.
It is also possible for haploid selection to drive the invasion of a loosely linked neo-W. 
Take, for instance, selection directionally favouring $A$ in both diploid sexes and meiotic drive in males.
Figure \ref{fig:regionplot_2a}B then shows that ploidally-antagonisitic selection can allow both neo-W haplotypes to invade. 
%}

%Neo-W alleles, on the other hand, can invade an ancestral XY system under some conditions (given in detail in the appendix). 
%%The full characteristic polynomials are given in the appendix (equations \ref{Binvasion} and \ref{Ainvasion}).
%Briefly, neo-W-$A$ and/or neo-W-$a$ haplotypes can spread when rare in the absence of recombination ($\lambda_{ma}>1$ and/or $\lambda_{mA}>1$), depending on the ancestral sex-ratio and allele frequencies. 
%Below we discuss the main forces determining the spread of these neo-W haplotypes and the impact of recombination for the overall success of the neo-W.
%To simplify our discussion we first outline the potential effects of haploid selection and then consider diploid selection in its absence.

%\textcolor{blue}{This explanation is clearer than the one I put in the appendix, if we drop it here, maybe we should try and re-word the corresponding part of the appendix using this paragraph}
%Haploid selection impacts the spread of neo-W haplotypes in three ways.
%Firstly, the zygotic sex ratio becomes male biased ($\zeta<1/2$) when the $a$ allele (which is fixed on the Y) is favoured during competition among male gametes or by meiotic drive in males. 
%This facilitates the spread of a neo-W because neo-W alleles cause the zygotes that carry them to develop as the rarer, female, sex. 
%Secondly, haploid selection in males affects the diploid genotypes of females by altering the allele frequencies in the male gametes that female gametes pair with.
%For instance, because an epistatically dominant neo-W always causes its carrier to become female, it creates females who carry either $Y-a$ or $X$ genotypes from their father.
%Thus, because when there is a polymorphism the $X$ carries some non-zero frequency of $A$, haploid selection in males impacts the diploid genotypes of females (e.g., creating more $Aa$ females when drive in males favours $Y-a$).
%How this affects the spread of the neo-W then depends on diploid and haploid selection in females.
%Thirdly, female drive and gamete competition directly select on neo-W haplotypes.
%Drive for $A$ in females favours neo-W-$A$ haplotypes, at a cost to neo-W-$a$ haplotypes, and vice-versa when there is drive for $a$.
%The impact of this drive depends on how often XX and neo-W females are heterozygous.
%Competition among female gametes acts similarly, and depends on the frequency of $A$ on resident X chromosomes (e.g., competition among eggs has no affect on the initial spread of the neo-W-$A$ haplotype when $A$ is fixed on the X).
%Because haploid selection in females favours one neo-W haplotype at the expense of the other, recombination off the favoured background becomes more detrimental as it becomes more favoured. Thus higher rates of recombination between the neo-W and the selected locus, $R$, can lead to smaller leading eigenvalues when there is haploid selection in females.

%In the absence of haploid selection and with the $A$ allele is fixed on the X, it is possible for both neo-W haplotypes can spread ($\lambda_{mA}>1$ and $\lambda_{ma}>1$ in \ref{Binvasion}), and thus neo-W invasion can occur regardless of its linkage to the selected locus.
%Invasion does not occur with purely sexually-antagonistic selection (i.e., $a$ directionally favoured in males and $A$ directionally favoured in females) because the X is then already as specialized as possible on the female sex.
%However, if, for example, $AA$ individuals suffer a fitness cost in females, yet $A$ is fixed on the X due to strong overdominance in males, both neo-W-$A$ and neo-W-$a$ haplotypes spread because they produce fewer unfit $AA$ females and never experience counter-selection in males.
%This is true even for the neo-W-$A$ haplotype because it can pair with a $Y-a$ haplotype and still be female. 
%When both haplotypes can spread alone the rate of recombination between the neo-W and the selected locus, $R$, does cannot prevent invasion, and thus the system can evolve looser sex-linkage (e.g., the neo-W could arise on an autosome, $R=1/2$).
%Even when only one haplotype can spread, invasion can still occur up to some positive rate of recombination, $R>0$ (as long as equation \ref{eq:lambdasGen} is satisfied). 
%That looser sex-linkage can evolve is contrary to the conclusions of \cite{vanDoorn:2010hu}, who did not explicitly calculate invasion fitness under ancestrally tight sex-linkage. \textcolor{blue}{or, just didn't calculate the equilibria (A) an (B).}
%Similar scenarios have been shown to select for a modifier that increases recombination between the sexes \citep[green regions of Figure 2 in][]{Otto2014}.
%\textcolor{blue}{n.b., the recombination regions are only subsets mostly because you also generate a really unfit Y haplotype through recombination. Probably not worth mentioning.}
%
%In the absence of haploid selection it is also possible for a neo-W to invade when there is a stable polymorphism at the $\mathbf{A}$ locus on X chromosomes.
%For example, overdominance in males and strong directional selection for $a$ in females creates a scenario that favours the spread of both neo-W haplotypes at equilibrium ($\lambda_{mA}>1$ and $\lambda_{ma}>1$ in \ref{Ainvasion}), as both haplotypes bring more $a$ alleles into females and never experience counter-selection in males.
%Thus, as in the case of the $A$ being fixed on the X, looser sex-linkage can evolve with a polymorphic X (i.e., $\lambda>1$ with $R>0$) and this is expected under the same scenarios that select for a modifier that increases recombination between the sex chromosomes \citep[blue regions of Figure 2 in][]{Otto2014}.

\subsubsection*{Loose linkage with the ancestral sex-determining region}

%\textcolor{blue}{As with the discussion, this section might need some editing for (1) unclearly describing looser linkage never evolving without haploid selection, which we now show is possible above and (2) going on about sex ratio selection too much - this should be changed to something along the lines of  ``male biased zygotic sex ratios ($\zeta-1/2$) are of the order of $\epsilon^2$. Therefore Fisherian sex ratio selection directly favours neo-Ws by a similar magnitude. 
%Perhaps surprisingly, neo-Ws can be favoured to the same degree by haploid selection in females when female haploid selection favours the alleles that tend to be carried by the neo-W.
%Thus XY to ZW is the same as ZW to XY even though sex ratio biases only exist in one case, given the assumptions made in this section.
%''
%}

%weak selection
%We are particularly concerned with whether or not a rare neo-sex-determining allele increases in frequency, which occurs when the largest eigenvalue, $\lambda$, is greater than one. 
%Here, we explicitly determine the conditions under which invasion occurs by assuming that the $A$ allele reaches an equilibrium frequency under the ancestral sex-determination system before the neo-sex-determination system ($m$) arises. 
%The equilibrium frequency of $A$ on different ancestral backgrounds ($\hat{p}^\male_Y$, $\hat{p}^\male_X$, and $\hat{p}^\female_X$) is given by equations \eqref{eq:pAve} and \eqref{eq:freq_diffs} where we assume selection and meiotic drive are weak relative to recombination ($s^\Hermaphrodite$, $t^\Hermaphrodite$, $\alpha_{\Delta}^\Hermaphrodite$ of order $\epsilon$). 
Assuming that selection is weak relative to all recombination rates ($r$, $R$ and $\chi$), we denote the leading eigenvalues describing the invasion of a neo-Y ($k=0$) and a neo-W ($k=1$) into an ancestrally XY system by $\lambda_{Y',XY}$ and $\lambda_{W',XY}$, respectively, which are

\begin{equation}
\lambda_{Y',XY} =1+ V_{A}S_{A}^2\frac{ \left( r-R \right) }{r R}+O\left(\epsilon^3 \right) 
\label{eq:lambda_neoY}
\end{equation}

\noindent 
and

\begin{equation}
\lambda_{W',XY} =\lambda_{Y',XY}+\left(2\alpha_{\Delta}^\male-2\alpha_{\Delta}^\female+t^\male-t^\female \right) \left( \hat{p}^\male_Y-\hat{p}^\male_X \right)/2
+O\left(\epsilon^3 \right) \\
\label{eq:lambda_neoW}
\end{equation}

\noindent
where $V_{A}=\bar{p}(1-\bar{p})$ is the variance in the frequency of $A$ and $S_{A}=(D^\male +\alpha_{\Delta}^\male+t^\male) - (D^\female+\alpha_{\Delta}^\female+t^\female)$ describes sex differences in selection for the $A$ versus $a$ across diploid selection, meiosis, and gametic competition.
The diploid selection term, $D^\Hermaphrodite=\big{(} \bar{p}s^\Hermaphrodite+(1-\bar{p})h^\Hermaphrodite s^\Hermaphrodite\big{)} -\big{(} \bar{p}h^\Hermaphrodite s^\Hermaphrodite+(1-\bar{p}) \big{)}$, is the difference in fitness between $A$ and $a$ alleles in diploids of sex $\Hermaphrodite \in \{\female,\male\}$, where $\bar{p}$ is the leading-order probability of mating with an $A$-bearing gamete from the opposite sex (see Appendix).

The neo-sex-determining allele, $m$, will spread if $\lambda_{m,XY}>1$. 
Equation \eqref{eq:lambda_neoY} demonstrates that under weak selection a neo-Y will invade if and only if it is more closely linked to the selected locus than the ancestral sex-determining region (i.e., if $R<r$; note that $V_{A}$ and $S_{A}^2$ are strictly positive as long as \textbf{A} is polymorphic). 
This echoes our tight linkage results above and the results of \citet{vanDoorn:2007eu}, who considered diploid selection only and also found that homogametic transitions (XY to XY or ZW to ZW) can occur when the neo-sex-determining locus is more closely linked to a locus under sexually-antagonistic selection. 
%Because tighter linkage between the sex-determining region and a selected locus evolves during these transitions, stronger associations between selected alleles and diploid sexes can build up by selection (e.g., between male-beneficial alleles and the male-determining allele, Y).%, increasing diploid mean fitness (Figure \ref{fig:Combination_MeanFit}).

Equation \eqref{eq:lambda_neoW} shows that, in contrast to the tight linkage results of the previous section, with weak selection and no haploid selection ($t^\Hermaphrodite=\alpha^\Hermaphrodite_{\Delta}=0$), as considered by \citet{vanDoorn:2010hu}, the spread of a neo-W is equivalent to the spread of a neo-Y ($\lambda_{W',XY}=\lambda_{Y',XY}$), such that heterogametic transitions (XY to ZW or ZW to XY) can also occur only if the neo-sex-determining region is more closely linked to a locus under selection ($R<r$). 
However, if there is any haploid selection, the additional term in equation \eqref{eq:lambda_neoW} can be positive, which can allow, for example, neo-W invasion ($\lambda_{W',XY}>1$) even when the neo-sex-determining region is less closely linked to the selected locus ($R>r$). 
These transitions are unusual because, when $R>r$, associations that selection has built up between alleles more favourable in one sex and alleles that determine sex will be weakened. 
Mean diploid fitness therefore decreases during heterogametic transitions that create looser sex-linkage (Figure \ref{fig:Combination_MeanFit}B,D).

Equation \eqref{eq:lambda_neoW} shows that neo-W alleles can invade an XY system for a large number of selective regimes. 
To clarify the parameter space under which $\lambda_{W',XY}>1$, we consider several special cases. 
Firstly, if the \textbf{A} locus is unlinked to the ancestral sex-determining region ($r=1/2$), a more closely linked neo-W ($R<1/2$) can always invade because there is then no association between $A$ alleles and sex chromosomes, $\left( \hat{p}^\male_Y-\hat{p}^\male_X \right)=0$. 
The second term in equation \eqref{eq:lambda_neoW} then disappears and invasion depends only on the sign of $(r-R)$. 
Indeed, invasion typically occurs when the neo-W is more closely linked to the selected locus than the ancestral sex-determining region, under a variety of selective regimes (Figure \ref{fig:Combination_Centimorgans}).
Secondly, we can simplify cases where invasion occurs despite looser sex-linkage, $R>r$, using the special case where $R=1/2$ and $r<1/2$ (e.g., the selected locus is on the ancestral sex chromosome and the novel sex-determining locus arises on an autosome). 
In table \ref{tab:specialcases} we give the conditions where invasion occurs when we further assume that haploid selection only occurs in one sex (e.g., during male meiosis only) and dominance coefficients are equal in the two sexes, $h^\female=h^\male$. 
When there is no gametic competition and meiotic drive is in one sex only, an unlinked neo-W can invade as long as the same allele is favoured during diploid selection in males and females ($s^\female s^\male>0$, see Figure \ref{fig:Combination_Centimorgans}B). %, which is 50\% of the parameter space. 
When there is no meiotic drive and gametic competition occurs in one sex only, an unlinked neo-W can invade as long as the same allele is favoured in male and female diploid selection and there are sex differences in selection of one type (e.g., $s^\female(s^\male-s^\female)>0$, see Figure \ref{fig:Combination_Centimorgans}C,D). %, which is 25\% of the parameter space. 
These special cases indicate that neo-W invasion can occur for a relatively large fraction of parameter space, even if the neo-W uncouples the sex-determining locus from a locus under selection.

\begin{table}[ht]
\centering
\smallskip
\caption{Invasion conditions for unlinked neo-W ($R=1/2$, $r<1/2$) into ancestral XY with one form of haploid selection}
\begin{tabular}{l l c }
\hline\hline
Scenario &  Assumptions & neo-W spreads ($\lambda_{W',XY}>1$) if \\ [0.5ex] \hline
%  $r=1/2$, $R<1/2$ & Always \\ [0.5ex] \hline
%  $r<1/2$, $R=1/2$ &  \\ [0.5ex] \hline
  male drive only & $h^\male=h^\female$, $t^\female=t^\male=\alpha^\female_{\Delta}=0$ & $s^\female s^\male>0$ \\
 female drive only & $h^\male=h^\female$, $t^\female=t^\male=\alpha^\male_{\Delta}=0$ & $s^\female s^\male>0$ \\
 sperm competition only &  $h^\male=h^\female$, $t^\female=\alpha^\female_{\Delta}=\alpha^\male_{\Delta}=0$ & $s^\female(s^\male-s^\female)>0$ \\
  egg competition only & $h^\male=h^\female$, $t^\male=\alpha^\female_{\Delta}=\alpha^\male_{\Delta}=0$ & $s^\male(s^\female-s^\male)>0$ \\
  \hline \hline
  \label{tab:specialcases}
 \end{tabular}
\end{table}

%If we further assume equal dominance ($h^\female=h^\male$) and that haploid selection only occurs during male meiosis ($t^\female=t^\male=\alpha^\female_{\Delta}=0$), or female meiosis ($t^\female=t^\male=\alpha^\male_{\Delta}=0$), neo-W invasion occurs when $s^\female s^\male>0$, Figure \ref{fig:Combination_Centimorgans}B.
%That is, when the same allele is favoured by selection in both males and females (it may remain polymorphic via counter selection during meiotic drive).

%\textcolor{red}{It may seem counterintuitive that, if the $A$ allele is more common on the ancestral-Y than the ancestral-X, and only favoured during haploid selection in males}

%\textcolor{red}{although our predictions also perform well when recombination is small, see figure \ref{fig:WinvasionLowR}. We would have to add a line showing what invasion fitness the weak selection approximation would give and add dots to this figure (dots that are currently connected by a line), as in (vD\&K, 2010) }

%Sex ratios
%Previous research suggests that when the ancestral sex-determining locus is linked to a locus that experiences haploid selection (e.g., meiotic drive), a new, unlinked sex-determining locus invades in order to restore equal sex ratios \citep{Kozielska:2010vm}. 
%Our model provides a good opportunity to determine whether Fisherian sex-ratio selection provides a useful explanation for the evolution of new sex-determining loci in other contexts. 
%Consider, for example, the case where the \textbf{A} locus is linked to the ancestral-SDR ($r<1/2$) and experiences meiotic drive in males only (e.g., during spermatogenesis but not during oogenesis, $\alpha^\male \neq 1/2$, $\alpha^\female=1/2$). 
%We will also disregard gametic competition ($t^\female=t^\male=0$) such that zygotic sex ratios are only biased by meiotic drive in males. 
%In this case, the zygotic sex ratio can be initially biased only if the ancestral sex-determining system is XY (Figure \ref{fig:Combination_Turnover}B). 
%If the ancestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Figure \ref{fig:Combination_Turnover}D).
%Thus, if the zygotic sex ratio is crucial to the evolution of new genetic sex-determining systems, invasion into ZW and XY systems will be distinct. 
%However, under weak selection we find that invasion by a homogametic neo-sex-determining allele (XY to XY or ZW to ZW) or by a heterogametic neo-sex-determining allele (XY to ZW or ZW to XY) occur under the same conditions. 
%That is, we can show that $\lambda_{Y',XY}=\lambda_{W',ZW}$ and $\lambda_{Y',ZW}=\lambda_{W',XY}$ (at least up to order $\epsilon^3$; for a numerical example, compare Figure \ref{fig:Combination_Turnover}A,B to Figure \ref{fig:Combination_Turnover}C,D).
%As it turns out, under weak selection the strength of sex-ratio selection favouring, say, the invasion of a neo-W in an XY system is the same as the strength of meiotic drive favouring the invasion of a neo-Y in a ZW system. 
%Even when these forces are not exactly the same (e.g., under tight sex-linkage; compare black and red curves near -25 and 25cM in Figure \ref{fig:Combination_Centimorgans}), it is important to remember that sex-ratio selection is only one of many potential selective forces acting to determine transitions between sex-determining systems.
%It is even possible for the other selective forces to overwhelm sex-ratio selection and favour sex-determination transitions that create sex-ratio biases (Figure \ref{fig:Combination_Turnover}A,C).

Previous research suggests that when the ancestral sex-determining locus is linked to a locus that experiences haploid selection (e.g., meiotic drive), a new, unlinked sex-determining locus invades in order to restore equal sex ratios \citep{Kozielska:2010vm}. 
Consider, for example, the case where the \textbf{A} locus is linked to the ancestral-SDR ($r<1/2$) and experiences meiotic drive in males only (e.g., during spermatogenesis but not during oogenesis, $\alpha^\male \neq 1/2$, $\alpha^\female=1/2$). 
Disregarding gametic competition ($t^\female=t^\male=0$) such that zygotic sex ratios are only biased by meiotic drive in males. 
In this case, the zygotic sex ratio can be initially biased only if the ancestral sex-determining system is XY (Figure \ref{fig:Combination_Turnover}B). 
%If the ancxestral sex-determining system is ZW, the zygotic sex ratio will be 1:1 because diploid sex is determined by the proportion of Z-bearing versus W-bearing eggs and meiosis in females is fair (Figure \ref{fig:Combination_Turnover}D).
We might therefore expect a difference in the potential for XY to ZW and ZW to XY transitions. 
However, to leading order with selection weak relative to recombination, we find that sex ratio selection (first terms in table \ref{tab:haplotype_growth}) is equal in magnitude to the fitness effects of alleles associated with new sex-determining alleles (second terms in table \ref{tab:haplotype_growth}).
Thus, invasion by a neo-W into an XY system and invasion by a neo-Y into a ZW system occur under the same conditions ($\lambda_{Y',XY}=\lambda_{W',ZW}$ and $\lambda_{Y',ZW}=\lambda_{W',XY}$, at least up to order $\epsilon^2$).
As selection becomes stronger (or linkage becomes tighter), this symmetry between sex-ratio selection and haploid selection is lost, causing differences in the strength of selection favouring the two heterogametic transitions (compare red to black near -25cM and 25 cM in Figure \ref{fig:Combination_Centimorgans}).

%Therefore Fisherian sex ratio selection directly favours neo-Ws by a similar magnitude. 
%Perhaps surprisingly, neo-Ws can be favoured to the same degree by haploid selection in females when female haploid selection favours the alleles that tend to be carried by the neo-W.
%Thus XY to ZW is the same as ZW to XY even though sex ratio biases only exist in one case, given the assumptions made in this section

\subsubsection*{Environmental sex determination}

We next consider the case where the new sex-determining mutation, $m$, causes sex to be determined probabilistically or by heterogeneous environmental conditions (environmental sex determiner, ESD). 
We assume that individuals carrying the $m$ allele develop as females with probability $k$ (e.g., in a fraction $k$ of the environments they randomly experience).
The characteristic polynomial determining the eigenvalues of the 8 equation system (equations \ref{eq:recursions}) does not reduce for ESD mutants as it does for $k=0$ or $k=1$. 
We therefore focus on weak selection here. 
Assuming weak selection, the spread of these mutations is given by

\begin{equation}
\begin{split}
\lambda_{ESD',XY} =& 1 + (1-2k)^2V_{A}{S_{A}}^2\frac{r-R}{r R} \\
&+\frac{k(\hat{p}^\male_Y-\hat{p}^\male_X)}{2}\left(k\left(2\alpha_{\Delta}^\male-2\alpha_{\Delta}^\female+t^\male-t^\female \right) -4(1-k)S_{A}\right)+O\left(\epsilon^3\right),
\end{split}
\label{eq:lambda_ESD_k}
\end{equation}

\noindent
which reduces to $\lambda_{Y',XY}$ when $k=0$ and $\lambda_{W',XY}$ when $k=1$.

%Under Fisherian sex-ratio selection, autosomal modifiers favour equal investment in male and female offspring, i.e., a 1:1 sex ratio \citep{Fisher:1930wy,Charnov:1982wg,West:2009we}. 
%A novel environmental sex-determiner that causes half of its carriers to become female and half to become male ($k=1/2$) will be in males half of the time and in females half of the time (like an autosome).
%In addition, these novel sex-determination alleles equalize the sex ratio and therefore one might expect them to be favoured by Fisherian sex-ratio selection when the resident sex ratio is biased.
%However, assuming weak selection, we find that the growth rate of a rare, dominant offspring-controlled neo-ESD allele that produces males or females with equal probability ($k=1/2$) is
Of particular interest are ESD mutations that cause half of their carriers to develop as females and half as males ($k=1/2$, creating equal sex ratios), the spread of which is given by
\begin{equation}
\lambda_{ESD',XY} =1+ \frac{1}{2}\frac{(\lambda_{Y',XY}-1) + (\lambda_{W',XY}-1)}{2} \Big|_{R=1/2} + O\left(\epsilon^3\right),
%V_A \frac{r}{1-2r}  \frac{s_{f} \left( s^\male - 3 s^\female\right) }{8 \left( s_{f} + s_{m} \right)^2} t_{m}^2.
\label{eq:lambda_ESD}
\end{equation}
\noindent
where we have indicated that $\lambda_{Y',XY}$ and $\lambda_{W',XY}$ are evaluated at $R=1/2$.
That is, recombination between the selected locus and the novel sex-determining locus, $R$, doesn't enter into the $k=1/2$ results.
This is because sex is essentially randomized each generation, %as if the novel sex-determiner were on an unlinked site that half the time acts as a neo-$W$ and half the time as a neo-$Y$.
preventing associations from building up between allele $A$ and sex.  
An important result from equation \eqref{eq:lambda_ESD} is that ESD can invade if there is haploid selection. 
When evaluated at $R=1/2$, $\lambda_{Y',XY} \leq 1$ but $\lambda_{W',XY}$ can be greater than one if there is haploid selection, as discussed above. 
Previous studies where ESD is favoured have typically assumed that environmental conditions (e.g., maternal condition, mate quality, age, or host size) can differentially affect the fitness of males versus females such that ESD invades because it allows sex determination to depend on the environment \citep[reviewed in][]{Charnov:1982wg,Bull:1983vi,West:2009we}. 
Here, ESD mutations can spread because they generate females that are either rare or have high fitness, in the same manner as a neo-W.

Equation \eqref{eq:lambda_ESD} also shows that invasion by a novel `perfect' ESD (equal sex ratio, $k=1/2$) mutation is the same for an ancestrally XY or ZW system (since $\lambda_{Y',XY} = \lambda_{W',ZW}$, $\lambda_{W',XY} = \lambda_{Y',ZW}$).
Thus, by the same argument as above (if drive only occurs in males then the sex ratio is only biased when the ancestral sex-determination system is XY), Fisherian sex-ratio selection alone does not explain the invasion of an offspring-controlled neo-ESD allele under weak selection. 
Rather, the neo-ESD gets half of the fitness of a feminizing mutation (neo-$W$) and half of the fitness of a masculinizing mutation (neo-$Y$), but only has an effect one half of the time (the other half of the time it produces the same sex as the ancestral system would have, to leading order). 
The net result can be that perfect ESD will not invade, even if current sex ratios are biased. 
For example, if there is haploid selection in males (either drive or pollen/sperm competition) but the conditions in table \ref{tab:specialcases} are not met, perfect ESD will not invade, even though it would equalize the zygotic sex ratio from an initially biased case (assuming $r<1/2$).

%Depends 50\% on its fitness relative to non-mutant males and 50\% on its fitness relative to non-mutant females.

%\subsubsection*{Maternally-controlled neo-ESD}

%Fisherian sex-ratio selection is sometimes considered in terms of balancing parental investment in male versus female offspring \citep{Charnov:1982wg}.
%In addition, under environmental sex-determination, the proportion of males/females is sometimes controlled by the mother (e.g., the proportion of eggs laid in warm versus cold environments). 
%We therefore also considered the invasion of a neo-sex-determining allele ($m$) in a model in which mothers that have at least one $m$ allele produce daughters with probability $k$. 
%As with offspring-controlled ESD, for all $k\in\{0,1/2,1\}$, we find that invasion into an ancestral XY system is the same as invasion into an ancestrally ZW system (at least up to order $\epsilon^3$, assuming weak selection), implying that transitions between genetic sex-determination and maternally controlled environmental sex-determination are not driven by Fisherian sex-ratio selection alone.
%\textcolor{red}{(Maternal ESD analysis still lacks meiotic drive -- Mathematica can't seem to deal with the added complexity.)}

%One might think that when the sex of zygotes is under the control of mothers, there would be strong selection to balance the sex ratio among zygotes. 
%However, we find that, with no meiotic drive and weak selection, the invasion fitness of a dominant, maternally controlled sex-determiner that produces proportion $k$ daughters can be written
%\begin{equation}
%\lambda_{k} = 1 + V_A S_A C_k + O\left(\epsilon^3\right),
%\end{equation}
%where
%\begin{equation}
%V_A = \frac{- \left[ t_m + h (s_{f}+s_{m}) \right] \left[ t_m + (1-h) (s_{f}+s_{m}) \right]}{(1-2h)^2(s_{f}+s_{m})^2}
%\end{equation}
%is the variance in $A$ ($V_A>0$ when the polymorphic resident equilibrium is stable), 
%with $h$ the dominance coefficient (equal in both sexes), 
%$C_k$ is a term that depends on $k$.
%
%%k=0
%\begin{equation}
%C_0 = big \; ugly \; term
%\end{equation}
%
%%k=1
%\begin{equation}
%C_1 = \frac{\left[ (2r(1-R)-R) s^\female + (1-2r)R s^\male \right]}{2}
%\end{equation}
%
%%k=1/2
%\begin{equation}
%C_{1/2} = even \; bigger, \; uglier \; term
%\end{equation}
%
%When $k=1$ the maternal- and offspring-controlled results are the same, $C_1 = C_W$, since the mutation remains in females (and hence mutant daughters always have mutant mothers).
%However, with $k\neq1$ the results with maternal control differ from those with offspring control.
%Interestingly, for all $k\in\{0,1/2,1\}$, we find that invasion into an ancestrally XY system is the same as invasion into an ancestrally ZW system (at least up to order $\epsilon^3$), implying that sex-ratio selection does not drive transitions between genetic sex determination and maternally controlled environmental sex determination.
%Of particular interest is $k=1/2$ (i.e., when the mother perfectly balances the sex ratio of her offspring).
%When both recombination rates are small we have $C_{1/2} \approx R(s^\male - s^\female)/8 = \lim_{r \rightarrow 0} C_1/4$.
%This implies that, at least under tight linkage, the invasion of maternally-controlled ESD is independent of $R$ (because $S_A\propto R^{-1}$) and can invade whenever a neo-$W$ can (which can invade even when it biases the sex ratio further; Figures \ref{fig:WinvasionLowR} -- \ref{fig:WinvasionHighR}).

\section*{Discussion}

%\textcolor{blue}{maybe re-order to put the results up front and focus on sex ratio less. Somewhere around these first two paragraphs we should include our tight linkage results. At the moment it doesn't make total sense because it assumes that tighter linkage is always favoured unless there is haploid selection.}

%MAIN RESULT
Two predominant theories explaining the remarkably high frequency of transitions between sex-determination systems are sexually-antagonistic selection and sex-ratio selection \citep[reviewed in][]{Blaser2012} \textcolor{red}{(van Doorn, 2014, sexual development)}.
The former predicts that neo-sex-determining alleles can invade when they arise in closer linkage with a sexually-antagonistic locus \citep{vanDoorn:2007eu,vanDoorn:2010hu}.
The latter predicts that neo-W alleles will invade an XY system when there is a male bias caused by haploid selection in males, and vice-versa, a neo-Y will invade a ZW system when there is a female bias caused by haploid selection in females \citep{Kozielska:2010vm,Ubeda:2015fx}.
Here we have shown that both predictions must be amended when recombination is weak relative to selection or selection happens in both diploid and haploid phases.

%TIGHT LINKAGE RESULTS
When the rate of recombination between the ancestral sex-determining locus and a locus under selection is small relative to the strength of selection (i.e., sex-linkage is tight, or selection is strong), heterogametic transitions (XY to ZW or ZW to XY) that reduce sex-linkage are possible, with or without haploid selection or sexually-antagonistic selection (Figure \ref{fig:regionplot_2a}).
The likelihoods of these transitions are driven by sex-ratio selection, direct selection on alleles linked to the neo-sex-determining allele, the ability of the neo-sex-determining allele to avoid selection in one sex, and the ability of the neo-sex-determining allele to bring alleles on the sex-specific chromosome in the ancestor into the other sex (given that the neo-sex determining allele is epistatically dominant to its predecessor).
This possibility that looser sex-linkage could evolve, even in the absence of haploid selection (Figure \ref{fig:regionplot_2a}A), was overlooked in \cite{vanDoorn:2010hu}, likely because they did not explicitly calculate the resident equilibria \citep[equation \ref{eq:tightequil};][]{Lloyd1977,Otto2014}.
Interestingly, there is substantial overlap between the parameter space that allows both neo-W-$A$ and neo-W-$a$ haplotypes to spread in an XY system and that which selects for increased recombination between X and Y chromosomes \citep[e.g., compare gray region of Figure \ref{fig:regionplot_2a}A with coloured regions of Figure 2(a) in][]{Otto2014}.
This makes sense, as when both neo-W haplotypes can spread the neo-W can invade despite reducing sex-linkage, i.e., the rate of recombination between the sex-determining allele and the selected locus increases.
%\textcolor{red}{explain difference too? also explain why a neo-Y doesn't invade to reduce linkage despite selection for more recombination? mention something about haploid selection (Fig7B)?} \textcolor{blue}{maybe leave for now}

%HAPLOID SELECTION CAN ACT LIKE SEXUALLY-ANTAGONISTIC SELECTION
%It has previously been demonstrated that new sex-determining systems can evolve if there is genetic variation maintained by sexually-antagonistic selection \citep{vanDoorn:2007eu,vanDoorn:2010hu}. 
%In particular, 
Under weak selection (or loose sex-linkage), transitions to new sex-determining systems can occur when they arise more closely linked to a sexually-antagonistic locus \citep{vanDoorn:2007eu,vanDoorn:2010hu}.
Our results show that genetic variation at loci that experience haploid selection can generate selection in favour of new sex-determining systems in a similar way. 
New sex-determining alleles are again favoured if they are more closely linked to a locus under haploid selection.
%HAPLOID SELECTION ALSO ALLOWS LOOSER LINKAGE TO EVOLVE
However, with haploid selection, heterogametic transitions (XY to ZW or ZW to XY) can also occur when the new sex-determining region is less closely linked to the locus under selection. 
Neo-W (neo-Y) alleles invade when their fitness in females (males) is greater than the mean fitness of females (males) under the ancestral sex-determination system and/or females (males) are the rarer sex.
With sexually-antagonistic selection (between diploid sexes) only, linkage between a selected locus and the sex-determining region strengthens associations between male beneficial alleles and the male-determining allele (Y or Z) and between female beneficial alleles and the female-determining allele (X or W). 
Thus, the mean fitness of both males and females increases with closer linkage to the sex-determining region. 
Therefore, new sex-determining alleles only invade if they are more closely linked than the ancestral sex-determining region. 
However, if there is haploid selection on loci linked to an XY (ZW) sex-determining region, selection can maintain polymorphisms at which the product of the frequency of females (males) and the mean fitness of females (males) is lower than it would be without sex-linkage. 
In these cases, unlinked neo-W (neo-Y) alleles can increase the frequency and/or fitness of the only sex they are found in, at a cost to the other sex, and invade despite lowering population mean fitness (Figure \ref{fig:Combination_MeanFit}). 
%\textcolor{red}{This is similar to mitochondria and male disease ...?}

%%SEX-RATIO SELECTION
%\textcolor{blue}{maybe drop paragraph?}
%\textcolor{red}{I messed with the sex-ratio selection paragraphs to tone down our "it doesn't matter" speech from before. Have at any amendments you'd like to make.}
%Linkage between haploid selected loci and sex-determining regions cases biased zygotic sex ratios \citep{Hamilton:1967ts,Burt:2006,Field:2012fd,Field:2013cc}.
%One might then expect Fisherian sex-ratio selection to drive the spread of new sex-determining systems that bring the sex ratio closer to 50:50. 
%Fisherian sex-ratio selection follows from the fact that, for an autosomal locus, half of the genetic material is inherited from a male, and half from a female \citep{Fisher:1930wy,West:2009we}. 
%Thus, if the population sex ratio is biased towards females, the average per-individual contribution of genetic material to the next generation from males is greater than the contribution from females (and vice versa for male-biased sex ratios). 
%Therefore, a mutant that increases investment in males will spread via the higher per-individual contributions made by males. 
%%%That is, under Fisherian sex-ratio selection, the success of a mutant relative to the non-mutant depends, in equal parts, on the contributions made by males and females to the next generation. 
%An implicit assumption of Fisherian sex-ratio selection is that the mutant allele is autosomal and has the same inheritance pattern as the non-mutant allele. 
%The mutations we consider here, neo-sex-determining alleles, break this assumption. 
%For example, the success of neo-Y/neo-W mutations depends only on the number of alleles contributed by males/females (Table \ref{tab:haplotype_growth}). 
%In this respect, a neo-W is similar to a cytoplasmic element, which also does not experience selection to balance sex ratios \citep{Frank:1989vl,Werren:1998co,Chase:2007bf}.
%Even mutants that are equally likely to be found in males or females, such as an environmental sex determination mutation (equation \ref{eq:lambda_ESD}), are not strictly autosomal if they determine sex. 
%Thus, despite the fact that sex ratio biases caused by gametic competition or meiotic drive have been shown to exert Fisherian sex-ratio selection on various autosomal modifiers \citep{Stalker:1961th,Smith:1975ft,Frank:1989vl,Hough:2013uo,Ubeda:2015fx, Otto:2015va}, we do not find evidence of Fisherian sex-ratio selection acting during invasion by neo-sex-determination systems (e.g., see Figure \ref{fig:Combination_Turnover} and \citealt{Ubeda:2015fx}, in which a neo-Y invades despite biasing sex ratios).

Sex ratio biases caused by gametic competition or meiotic drive have been shown to exert Fisherian sex-ratio selection on various autosomal \citep{Stalker:1961th,Smith:1975ft,Frank:1989vl,Hough:2013uo,Ubeda:2015fx, Otto:2015va} and sex-linked \citep{Ubeda:2015fx} modifiers.
We find that sex-ratio biases caused by haploid selection can also affect transitions between sex-determining systems (e.g., see $\zeta$ terms in Table \ref{tab:haplotype_growth}). 
For instance, when an allele that drives in males is linked to an XY locus it will often become associated with the Y and therefore produce a male bias ($\zeta<1/2$).
This male bias increases the potential for a neo-W to invade (as we then have $(2\zeta)^{-1}>1$ in Table \ref{tab:haplotype_growth}), which can equalize the sex-ratio \citep[for a related example see][]{Ubeda:2015fx}.
%However, the neo-W may not equalize the sex-ratio (e.g., if the driving allele also drives in females or affects competition among eggs)
However, this sex-ratio selection can be overwhelmed when the driving allele has additional selective effects (e.g., when it is beneficial for male diploids but detrimental for female diploids; Table \ref{tab:specialcases}), preventing the neo-W from invading. %and equalizing the sex ratio.
Indeed, these additional selective effects can even favour transitions between sex-determining systems that create new sex-ratio biases.
For example, in an ancestral ZW system, an allele that drives only in males can allow a linked neo-Y to invade, despite the fact it creates a male bias (Figure \ref{fig:Combination_Turnover}C).
%\textcolor{blue}{and figure 1, where the end of one panel is the same as the beginning of another, however, this figure may change).}
%What we would like to stress is that sex-ratio selection alone cannot predict when new sex-determining systems can evolve.
Furthermore, with weak selection, there is no asymmetry between XY to ZW and ZW to XY transitions, indicating that sex-ratio selection does not dominate (i.e., the sex-ratio bias created by haploid selection impacts the spread of a neo-W into an XY system the same way it impacts the spread of a neo-Y into a ZW system with a 1:1 sex ratio). 
An asymmetry can develop when sex-linkage is tight (e.g., Figure \ref{fig:Combination_Centimorgans} near -25cM and 25cM) but under most circumstances we do not predict asymmetry between XY to ZW and ZW to XY transitions despite the presence/absence of sex ratio selection. 
Thus, haploid selection can favour heterogametic transitions both via sex-ratio selection and via fitness effects of alleles that are associated with the neo-sex-determining allele, and these selection pressures are often of equal magnitude. 
%\textcolor{blue}{nice, pretty subtle argument but I think that's a good job.}

%SEX-RATIO BIAS IN UBEDA ET AL
%DRAFT (improve): In \citet{Ubeda:2015fx}, the new sex determining locus spreads because it arises in linkage with a locus that experiences drive. They assume that drive occurs predominantly in one sex, e.g., during spermatogenesis or a 'killer' sperm. A driving allele is maintained at an intermediate frequency by selection, e.g., because it causes male sterility when homozygous (because all male sperm are killed). Y chromosomes that arise in linkage with the driving allele spread because they allow drive to occur more often, thus genetic sex determination with a sex ratio bias evolves. 
%Thus \citet{Ubeda:2015fx} also find that genetic sex determiners can invade, despite causing sex ratios to become biased. 
%Finally, they show that autosomal 'restorers' that negate the effects of meiotic drive can invade and restore an equal sex ratio.

%WHY SEX-RATIO BIASES CAN BE FAVOURED
%We note two other ways in which sex determination has been shown to relate to zygotic sex ratios.
%Firstly, female-biased sex ratios can be favoured when there is local mate competition, where all matings are between siblings and assuming one male can inseminate many females \citep{Hamilton:1967ts}. 
%Therefore, with local mate competition, feminizing mutations can spread because they bias the sex ratio towards females \citep{Wilson:1981vm,Vuillleumier:2007bh}. 
%Secondly, environmental conditions (e.g., maternal condition, mate quality, age, or host size) can differentially affect the fitness of males versus females such that the optimal allocation to males/females depends on the environment \citep{Trivers:1973wb,Charnov:1977tx,Charnov:1982wg}. 
%In such cases, flexible sex-determination systems may evolve in order to allow the zygotic sex ratio to be determined in a way that depends on the environment \citep{Charnov:1977tx,Werren:1984tl,Pen:2010kk}. 
%In this study, we do not consider environmental condition dependence or local mate competition \citep[reviewed in][]{Charnov:1982wg,Bull:1983vi,West:2009we}.

%Limitation 1: DEGENERATE Y
We assume that sex-determining alleles do not experience direct selection except via their associations with sex and alleles at a selected locus. 
However, in some cases, there may be significant degeneration around the sex-limited allele (Y or W) in the ancestral sex-determining region because recessive deleterious mutations and/or deletions may fix around the Y or W allele \citep{Rice:1996ke,Charlesworth:2000cc,Bachtrog:2006ed,Marais:2008hm}. 
During heterogametic transitions (XY to ZW or ZW to XY), the formally sex-limited allele fixes such that all individuals have YY or WW genotypes (Figure \ref{fig:Combination_Turnover}). 
Any recessive deleterious alleles linked to the Y or W will therefore be revealed to selection during a heterogametic transition. 
This phenomenon was studied by \citet{vanDoorn:2010hu}, who found that degeneration can prevent fixation of a neo-W or a neo-Y allele, leading to a mixed sex-determination system where the ancestral- and neo- sex-determining loci are both polymorphic. 
However, they noted that very rare recombination events around the ancestral sex-determining region can allow these heterogametic transitions to complete.  
While not explicitly studied, we also predict that Y or W degeneration would prevent fixation of the new sex-determiners considered here.

%Limitation 2: TWO ALLELE DRIVE
In addition, our model of meiotic drive is simple, involving a single locus with two alleles. 
However, many meiotic drive systems involve an interaction with another locus at which alleles may `suppress' the action of meiotic drive \citep[][]{Burt:2006,Lindholm:2016cw}.
Thus, the dynamics of meiotic drive alleles can be heavily dependent on the interaction between two loci and the recombination rate between them, which in turn can be affected by sex-linkage if there is reduced recombination between sex chromosomes \citep{Hurst:1991uh}.
Furthermore, in some cases, a driving allele may act by killing any gametes that carry a `target' allele at another locus, in which case there is a two-locus drive system and the total number of gametes produced can be reduced by meiotic drive. 
Where gamete number is reduced by meiotic drive, the number of mates competing for fertilization (mating system) can affect the equilibrium frequency of a meiotic drive allele \citep{Holman:2015en}. 
In polygamous mating systems, the intensity of pollen/sperm competition can depend on the density of males available to donate pollen/sperm, which can itself depend on the sex ratio \citep{Taylor:2002wu}. 
Since the sex ratio is partly determined by the sex-determination system, the evolution of new sex-determination system could by influenced by these dynamics. 
How the evolution of new sex-determining mechanisms could be influenced by two-locus meiotic drive and/or by ecological feedbacks under different mating systems remains to be studied.

%PREDICTIONS TO TEST
The hypotheses presented here can be empirically investigated in a similar manner to the idea that transitions between sex-determining systems are favoured by linkage to sexually-antagonistic variation. 
In the case of sexually-antagonistic variation, one supporting observation is that genes expected to be under sexually-antagonistic selection (e.g., those causing bright male colouration) have been found on recently derived sex chromosomes \citep{Lindholm:2002dw,Tripathi:2009cw,Ser:2010iq}. 
Our results suggest that polymorphic loci that are ancestrally sex-linked and under sex-specific selection could also drive heterogametic transitions between sex-determination systems. 
As noted by \citet{vanDoorn:2010hu}, it would be prudent to compare closely related clades in order to determine whether observed polymorphisms pre-dates a transition in sex-determination or arose afterwards, particularly because sex-linkage allows sexually-antagonistic selection to maintain polymorphisms under a different and larger parameter space \citep{Rice:1987hs,Jordan:2011fj}. 
As with sexually-antagonistic selection, the presence of haploid selected loci around ancestral- or novel-sex-determining regions could support their role in sex chromosome turnover.
%We note that linkage with sex chromosomes is not, a priori, more permissive to the maintenance of ploidally antagonistic variation \citep{Immler:2012tl}. 
A recent transcriptome analysis in \textit{Rumex}, suggests a role for haploid competition in the evolution of sex-determination systems by showing that Y-linked genes are overexpressed in pollen but not in male diploids, indicating variation currently or previously maintained by haploid selection; over-expression also occurs on the autosome that is orthologous to the sex chromosomes in closely related species \textcolor{red}{(Sandler et al., 2017, Personal Communication)}.
% (\textcolor{red}{George Sandler, a student in the Wright and Barrett labs, has done some yet-to-be-published work on Rumex that we should cite here. We can send him this draft and get his permission to cite him as personal communication or unpublished results. I think he has basically found that genes retained on the Y are overexpressed in pollen but not in male diploids, suggesting they are being maintained by haploid selection, not sexual antagonism. And I think they know that the sex chromosomes are recently derived, and that in closely related species the autosome matching up with the newly derived sex chromosome also has genes that are over expressed in pollen. I guess this is a follow up to Crowson et al 2017 Mol Biol Evol 34:1140, which we could potentially cite as well.}).

%HOW TO TEST EMPIRICALLY
%\textcolor{blue}{I think Sally was sceptical of this idea. i.e., if the second term in \ref{eq:lambda_neoW} is negative, then get less transitions with haploid selection included, right? }
%\textcolor{red}{Yeah, and given that only 1/2 the discovered plant sex chromosomes are hetermorphic and there are so few plant sex chromosomes to begin with (Charlesworth 2016 Annual Review Plant Biology), might it be the case that haploid selection is constraining things? perhaps we could think of some particular forms of haploid selection that might be common and give their predicted effects, or just say some things that are possible with haploid selection that are not without it -- if there are any. just looking at eq 3 though, its gonna be hard to get a low freq of A on the Y if haploid selection in males favours A, i.e., it should often be positive when there is no female haploid selection, right?}
%\textcolor{blue}{That sounds right to me and is backed up by equation \eqref{eq:freq_diffs}. It may be technically possible for that term to be negative though. Re: hetermorphism, I see that as a slightly different issue. It's thought that haploid selection may prevent deletions from becoming fixed on the Y, like in the Crowson (2017) paper. 
%I think the issue of `why so few plant sex chromosomes' is partly an issue of detection and also `why so much hermaphroditism'. Is it possible that the ESD results could show selective forces that also favour transitions from dioecy to hermaphroditism? I think so. Our ESD mutant produces 1/2 females and 1/2 males whereas a hermaphrodite invests roughly equally in each type of gamete produced. We don't consider the forces that are usually thought to be involved in these transitions though (inbreeding depression, reproductive assurance and possibly others?). }
%Taken at face value, our results indicate that transitions in heterogametey (XY to ZW or vice versa) are more likely to be favoured by selection if there is selection upon both haploid and diploid genotypes rather than diploid selection alone (). 
%This prediction could be examined using a suitable proxy for haploid selection, for example, \cite{Lenormand:2005vb} use the outcrossing rate in plants as a proxy for the strength of pollen competition. 
%In animals, one might expect gametic competition to be stronger in species where sperm is required to live for a long time after spermatogenesis because transcripts shared during spermatogenesis may become depleted, revealing the haploid phenotype of the sperm \citep{Immler:2014im}. 
%Given the caveats mentioned above about the form of meiotic drive modelled, we would also expect that heterogametic transitions in sex determination would be more common in clades where there is meiotic drive. 
Taken at face value, our results indicate that transitions in heterogametey (XY to ZW or vice versa) are more likely than transitions in homogamety when genetic conflict is predominately between the haploids of each sex (e.g., with $|D^\male - D^\female| << |\alpha_\Delta^\male - \alpha_\Delta^\female + t^\male - t^\female|$ we have $\lambda_{W',XY} > \lambda_{Y',XY}$; equations \ref{eq:lambda_neoW} and \ref{eq:freq_diffs}). 
In addition, because haploid selection can cause transitions that increase or decrease sex-linkage, haploid selection may lead to less stability, and greater potential for cycling, in sex-determination systems (e.g., the final state in Figure \ref{fig:Combination_Turnover}C is the starting state in Figure \ref{fig:Combination_Turnover}B). 
Potentially, successive heterogametic transitions between master regulators of sex-determination could be inferred from careful examination of the molecular pathways by which sex is determined. 
Our predictions could also be examined using a suitable proxy for haploid selection, for example, \cite{Lenormand:2005vb} use the outcrossing rate in plants as a proxy for the strength of pollen competition. 
Furthermore, assuming that transitions from dioecy to hermaphroditism (equal parental investment in male and female gametes) are favoured in a similar manner to the ESD examined here (equal probability of zygotes developing as males or females), our results suggest that haploid competition during the multicellular haploid stage could drive transitions between dioecy and hermaphroditism in plants \textcolor{red}{(Käfer et al., 2017, Sabath et al., 2017)}.  
In animals, one might expect gametic competition to be stronger in species where sperm is required to live for a long time after spermatogenesis because transcripts shared during spermatogenesis may become depleted, revealing the haploid phenotype of the sperm \citep{Immler:2014im}.
Given the caveats mentioned above about the form of meiotic drive modelled, we would also expect that heterogametic transitions in sex determination would be more common in clades where there is meiotic drive.

%\textcolor{blue}{similarly, this part doesn't discuss our new tight linkage results or our new perspective on sex ratios.}
%CONCLUSION
We have shown that tight sex-linkage and haploid selection can drive previously unexpected transitions between sex-determination systems.
In particular, both can select for neo-sex-determining loci that are more loosely linked. 
In addition, haploid selection alone can cause transitions analogous to those caused by purely sexually-antagonistic selection, eliminating the need for differences in selection between male and female diploids.
Perhaps counterintuitively, transitions involving haploid selection can be driven by sex-ratio selection, or cause sex-ratio biases to evolve. 
We therefore argue that haploid selection should be considered, alongside sexually-antagonistic and sex-ratio selection, as a potentially pivotal factor in the evolution of many sex-determination systems. 
Overall, our results suggest several new scenarios under which new sex-determination systems are favoured, including sex-specific selection on ancestrally sex-linked loci, which could help to explain why the evolution of sex-determination systems is so dynamic. 
%Further, we have shown the way in which transitions are affected by haploid selection is not intuitively obvious. 
%Firstly, sex-specific haploid selection affects turnovers between sex-determination systems in a manner that is qualitatively different from diploid sex-specific selection. 
%In particular, closer linkage between a sex-determining locus and a selected locus is not always favoured during heterogametic transitions when there is haploid selection. 
%Secondly, even though haploid selection is a source of zygotic sex-ratio biases, in our models Fisherian sex-ratio selection does not have good explanatory power in determining whether various sex-determination systems evolve. 
%This result is surprising given that sex ratios are ultimately determined via the sex-determination system, and leads us to the conclusion that three selective forces -- haploid, diploid, and sex-ratio selection -- should all be considered when exploring transitions between sex-determination systems.

%POLYGENIC SEX DETERMINATION
%\textcolor{red}{Numerical results suggest that polygenic sex determination can be stable with gametic competition. i.e., feminizers/masculinizers sometimes invade but don't fix. We have not discussed this. Might be interesting because it appears that polygenic sex determination exists (possibly stably, rather than transitionally) but previous models don't show any advantage to polygenic sex determination. There are two other ways that polygenic sex determination could be observed. (1) Direct selection against WW or YY genotypes during heterogametic transitions, mentioned above. (2) Loss of variation at the selected locus: The conditions under which polymorphism is maintained can depend on linkage to the SDR, and when a neo-SDR evolves you can move into a region where $A$ or $a$ fixes, at which point there is no longer any selection.}

%``Polygenic sex determination has been reported in many plants (e.g. Shannon \& Holsinger 2007), fishes (Vandeputte et al. 2007; Ser et al. 2010; Liew et al. 2012), crustaceans (e.g. Battaglia 1958; Battaglia \& Malesani 1959; Voordouw
%\& Anholt 2002), bivalves (Haley 1977; Saavedra et al. 1997), gastropods (Yusa 2007a,b), and polychaetes (Bacci 1965, 1978; Premoli et al. 1996).''
%From Vuilleumier et al. 2007:
%``Polymorphism for sex-determining genes within or among populations has been reported in many species including houseflies, midges, woodlice, platyfish, cichlid fish, and frogs (Gordon, 1944; Kallman, 1970; Thomp-son, 1971; Macdonald, 1978; Bull, 1983; Rigaud et al., 1997; Caubet et al., 2000; Lande et al., 2001; Ogataet al., 2003; Lee et al., 2004; Mank et al., 2006).''
%Also check Kallman (1984) -from vD\&K, 2010.

\bibliographystyle{amnatnat.bst}
\bibliography{sex_chromosomes.bib}

\newpage
\section*{Figures}
%\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Heterogametic transitions with/without sex ratio change
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\begin{tikzpicture}[remember picture]
	\node[] (plot) {\includegraphics[width=\linewidth]{Combination_Turnover}};
	\node[above = 0cm of plot, xshift = -70] () {$r=0.5$, $R=0.05$};
	\node[above = 0cm of plot, xshift = 120] () {$r=0.05$, $R=0.5$};
	\node[left = 0cm of plot, yshift = 105, rotate = 90] () {XY $\rightarrow$ ZW};
	\node[left = 0cm of plot, yshift = -20, rotate = 90] () {ZW $\rightarrow$ XY};
\end{tikzpicture}
\caption{
Heterogametic transitions from XY to ZW sex determination (neo-W frequency shown by black lines, panels A and B) or from ZW to XY (neo-Y frequency shown by black lines, panels C and D) occur similarly regardless of sex ratio biases present before (B versus D) or after (C versus A, dashed lines show male frequency). 
During invasion by a neo-ZW sex-determination system (A and B), the ancestral Y fixes in both males and females (blue and red lines). 
Similarly, the ancestral W allele fixes in males and females (blue and red lines) during a ZW to XY transition. 
In this plot, there is no gametic competition ($t^\female=t^\male=0$) and meiotic drive occurs during male meiosis only ($\alpha^\female_{\Delta}=0$, $\alpha^\male_{\Delta}=-1/5$). Therefore, sex ratio biases can only arise when the \textbf{A} locus is linked to an XY sex-determining locus.
In panels A and C, the neo-sex-determining locus is more closely linked to the \textbf{A} locus than the ancestral sex-detemining region ($r=1/2$, $R=1/20$) such that a neo-Y can caused biased sex ratios (panel C).
In panels B and D, the ancestral sex-determining locus is more closely linked to the \textbf{A} locus than the neo-sex-detemining locus ($r=1/20$, $R=1/2$). 
Therefore, an ancestral XY sex determination can have a biased zygotic sex ratio that becomes unbiased after an unlinked neo-W invades (B). 
However, in panel D, a unlinked neo-Y invades an ancestral ZW sex-determination system in a similar manner but no biases to the zygotic sex ratio occur. 
With diploid selection alone, neo-sex-determining loci do not spread if they are less closely linked to the \textbf{A} locus than the ancestral sex-determining locus (see equation \eqref{eq:lambda_neoW} and Figure \ref{fig:Combination_Centimorgans}A). 
In this plot there are no sex differences in selection and an equilibrium is maintained because selection in diploids opposes meiotic drive, $s^\female =s^\male = 1/5$, $h^\female = h^\male = 7/10$.
\textcolor{red}{Aesthetic adjustments: 
%Could add titles to the columns/rows: neo-W for row 1, neo-Y for row 3, $r=0.5$, $R=0.05$ for column 1 and $r=0.05$, $R=0.5$ for column 2. 
%Could adjust padding (too much whitespace where there is no axis label). 
%It also seems could increase ratio of font size relative to plot size to make figure more compact. 
%Matt - could you uncomment the line legends in the Mathematica file (function not included in my Mathematica version).
Add chromosome cartoons to depict recombination rates?
}
}
\label{fig:Combination_Turnover}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Heterogametic transitions with/without sex ratio change
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\begin{tikzpicture}[remember picture]
	\node[] (plot) {\includegraphics[width=\linewidth]{Combination_Turnover_Ratio}};
	\node[above = -1cm of plot, xshift = -100] () {XY $\rightarrow$ ZW};
	\node[above = -1cm of plot, xshift = 90] () {ZW $\rightarrow$ XY};
\end{tikzpicture}
\caption{
\textcolor{red}{Is this what Sally was thinking?}
\textcolor{blue}{I think this works but I'm confused as to why the speed of spread should be so much different for XY and ZW here. Figure \ref{fig:Combination_Turnover} and \ref{fig:Combination_Centimorgans} suggests that there's not much difference between XY->ZW and ZW->XY. Maybe we should just stick with the 4 cases in figure \ref{fig:Combination_Turnover}. }
\textcolor{red}{i messed up and had $r=0.5$ instead of $r=0.05$ for B and D. haven't fixed fitness plot below.}
}
\label{fig:Combination_Turnover_Ratio}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Changes in mean fitness
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\begin{tikzpicture}[remember picture]
	\node[] (plot) {\includegraphics[width=\linewidth]{Combination_Fitness_Rec}};
	\node[above = -1cm of plot, xshift = -100] () {XY $\rightarrow$ ZW};
	\node[above = -1cm of plot, xshift = 90] () {ZW $\rightarrow$ XY};
\end{tikzpicture}
\caption{
\textcolor{red}{This complicated thing matches the plot above. I don't think we want to include it...?}
}
\label{fig:Combination_Fitness_Rec}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Changes in mean fitness
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\begin{tikzpicture}[remember picture]
	\node[] (plot) {\includegraphics[width=\linewidth]{Combination_MeanFit}};
	\node[above = 0cm of plot, xshift = -70] () {$r=0.5$, $R=0.05$};
	\node[above = 0cm of plot, xshift = 120] () {$r=0.05$, $R=0.5$};
	\node[left = 0cm of plot, yshift = 105, rotate = 90] () {XY $\rightarrow$ ZW};
	\node[left = 0cm of plot, yshift = -20, rotate = 90] () {ZW $\rightarrow$ XY};
\end{tikzpicture}
\caption{
Changes in mean diploid fitness of males (blue lines), females (red lines), and the entire population (black lines) during the transitions between sex-determination systems shown in Figure \ref{fig:Combination_Turnover}. 
Here mean diploid fitness of a particular sex is its mean diploid viability fitness times twice its frequency in the population, to capture the fact that epistatically dominant sex-determining alleles can also invade because they selfishly make more of the sex they are in. 
The mean fitness of females increases during the spread of neo-W alleles (A and B) and the mean fitness of males increases during the spread of neo-Y alleles (C and D). 
However, when a neo-sex determining system evolves that is less closely linked to a locus under selection (B and D), population mean fitness decreases. 
%\textcolor{red}{
%Could add titles to the columns/rows: neo-W for row 1, neo-Y for row 3, $r=0.5$, $R=0.05$ for column 1 and $r=0.05$, $R=0.5$ for column 2. 
%%\& possibly adjust padding (too much whitespace?). 
%%Matt - could you uncomment the line legends in the Mathematica file (function not included in my Mathematica version).
%}
%\textcolor{red}{I'm still confused why male and female mean fitnesses aren't normalized by their frequency. I'm not sure we should be calling them means without this normalization step. Or we should justify this by saying that mean fitness also has something to do with the number of a sex, i.e., multiply *real* mean fitness in females by freqfemale/(1/2)? See the next figure for what happens when we do normalize.}
\textcolor{blue}{I think we should give this plot showing (male mean fitness * freq males) and (female mean fitness * freq females), without multiplying by 2 (leave off black lines, population mean fitness). 
We could also re-plot the sex ratios on this same scale. 
The plot below, `adjusted for sex ratio', could then go in the appendix. The point is that neo-W (neo-Y) can invade when the frequency of females (males) multiplied by their mean fitness increases.  }
}
\label{fig:Combination_MeanFit}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Changes in mean fitness corrected for sex ratio
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\begin{tikzpicture}[remember picture]
	\node[] (plot) {\includegraphics[width=\linewidth]{Combination_MeanFit_Corrected}};
\end{tikzpicture}
\caption{
\textcolor{red}{Last plot with mean fitness of sexes corrected for sex ratio. Could add to previous plot with dashed lines?}
}
\label{fig:Combination_MeanFit_Corrected}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Chromosomal Positions
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
%\includegraphics[width=\linewidth]{Combination_Centimorgan}\\
%\includegraphics[width=\linewidth]{PositionPlot}\\
\begin{tikzpicture}[remember picture]
%	\node[] () {\includegraphics[width=\linewidth]{PositionPlot}};
	\node[] () {\includegraphics[width=\linewidth]{PositionPlot_2}};
	\node[anchor=east] at (0, 1.5) {\footnotesize diploid sex-antagonism}; %, $s^\male s^\female < 0$};
	\node[anchor=east] at (6.5, 1.5) {\footnotesize haploid sex-antagonism}; %, $s^\male s^\female > 0$};
	\node[anchor=east] at (0, -3) {\footnotesize male drive (ploidy-antagonism)}; %, $s^\male < s^\female$};
%	\node[anchor=east] at (6.5, -3) {\footnotesize pollen competition (ploidy-antagonism)};%, $s^\female < s^\female$};
	\node[anchor=center] at (-1.3, 4.25) {\scriptsize tighter sex-linkage};
	\node[anchor=center] at (2.45, 3.5) {\scriptsize ancestral SDR};
	\draw[->,thick] (2.45, 3.25) -- (2.45, 2.5);
	\node[anchor=center] at (-1.4, -0.25) {\scriptsize novel SDR};
	\draw[->,thick] (-1.4, -0.5) -- (-1.4, -1);
\end{tikzpicture}
\caption{
Invasion fitness of a neo-sex-determining allele plotted against the relative genomic location of a locus under direct selection, \textbf{A}, for various selective regimes. 
The ancestral sex-determining locus is located at -25 and the novel sex-determining locus is located at 25. %and that there is a polymorphism at the \textbf{A} locus maintained by selection.
We use Haldane's map function \citep[Equation 3 in ][]{Haldane1919} to convert from map distance (centiMorgans) to the probability of a cross-over event. 
In A, there is no haploid selection ($t^\Hermaphrodite=\alpha^\Hermaphrodite_{\Delta}=0$) and selection in diploids is sexually antagonistic \citep[following][]{vanDoorn:2010hu}, in which case a neo-W can only invade if it is more closely linked to the selected locus ($s^\female=1/10$, $h^\female=7/10$ $s^\male=-1/10$, $h^\male=3/10$).
In B-D we include haploid selection. %and assume that selection in diploids is not sexually-antagonistic ($s^\female s^\male>0$). 
%A polymorphism can then be maintained by opposing selection between the haploid and diploid phases. 
%In B, there is drive in favour of the $a$ allele in males ($\alpha^\male_{\Delta}=-1/20$), no female meiotic drive or gametic competition, $t^\Hermaphrodite=\alpha^\female_{\Delta}=0$), and equal selection in diploid sexes ($s^\female=s^\male=1/10$, $h^\female=h^\male=7/10$). 
%In this case, a neo-W can invade even when the selected locus is more closely linked to the ancestral sex determining locus (see Table \ref{tab:specialcases} and Figure \ref{fig:Combination_Turnover}). 
%In C and D, there is gametic competition among male gametes only (favouring $a$, $t^\male=-1/10$) and no meiotic drive or gametic competition in females ($t^\female=\alpha^\Hermaphrodite_{\Delta}=0$). 
%In this case, the neo-W does not invade if $s^\female>s^\male$ (panel C: $s^\female=3/20$, $s^\male=1/20$) but does if $s^\female<s^\male$ (panel D: $s^\female=1/20$, $s^\male=3/20$), see Table \ref{tab:specialcases}.\\
In B
%\textcolor{red}{
%1. I suspect that panel C has a region where no equilibrium is maintained (CHECK! Maybe include different parameters here or remove the part when no equilibrium). MMO: If you trust the sieve function there are stable equilibria across the entire range, although they differ greatly between XY and ZW systems near -25cM.
%\\
%2. Currently use different parameters for B than using in figure \ref{fig:Combination_Turnover} (selection/drive twice as strong in turnover figure). MMO: this is to keep it within the bounds of the plot -- using the same parameters as figure 1 makes the peak at -25 reach roughly 0.1, and then it is difficult to see the details of A,C, and D.
%%This plot would also benefit from titles giving, e.g., ``sexually-antagonistic selection, $s^\female s^\male<0$'' for A, ``male meiotic drive, $s^\female s^\male>0$'' for B
%}
}
\label{fig:Combination_Centimorgans}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Haplotype invasion region plot
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!h]
\centering
\begin{tikzpicture}[remember picture]
	\node[] () {
	\includegraphics[width=0.5\linewidth]{regionplot_2a}
	\includegraphics[width=0.5\linewidth]{regionplot_hap}
	};
\end{tikzpicture}
\caption{
Parameter space (gray) where both neo-W haplotypes can invade from the same stable resident $r=0$ equilibria (equations \ref{eq:tightequil}), and therefore where an unlinked neo-W can invade an XY system with perfect sex-linkage.
\textbf{A}, In the absence of haploid selection, both neo-W haplotypes can invade for much of the parameter space where the relative fitnesses of male homozygotes, $w_{AA}^\male$ and $w_{aa}^\male$, are both less than that of the heterozygote, $w_{Aa}^\male = 1$.
In the white region neo-W haplotypes paired with the allele fixed on the Y cannot invade.
Parameters as in \cite{Otto2014} Figure 2a: $w_A^\Hermaphrodite=w_a^\Hermaphrodite$, $\alpha^\Hermaphrodite=1/2$, $w_{Aa}^\Hermaphrodite=1$, and $w_{AA}^\female=w_{aa}^\female=0.75$.
\textbf{B}, When selection is the same in both diploid sexes ($w_{aa}^\Hermaphrodite = 1$, $w_{Aa}^\Hermaphrodite = 1 + h s$, $w_{AA}^\Hermaphrodite = 1+s$), both neo-W haplotypes can invade over a portion of the parameter space where selection in diploids ($s$) opposes the force of drive during meiosis in males ($\alpha_{\Delta}^\male$).
Parameters: $w_A^\Hermaphrodite=w_a^\Hermaphrodite$, $\alpha^\female=1/2$, $h=1/2$.
}
\label{fig:regionplot_2a}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%W invades XY and fixes
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{FreqWLowR}\\
%\includegraphics[width=0.5\linewidth]{SexRatioLowR}\\
%\includegraphics[width=0.5\linewidth]{MeanDipFitLowR}
%\caption{
%Haploid selection allows a neo-$W$ to invade an ancestral $XY$ system and fix (\textbf{A}) despite temporarily biasing the sex ratio further (\textbf{B}) and decreasing mean diploid viability fitness (\textbf{C}).
%Complete turnover between genetic sex-determination systems occurs despite the neo-$W$ being less tightly linked to the selected locus than the ancestral sex-determining locus is, $R>r$.
%Parameters: $k=1$, $s^\female = 0.05$, $s^\male = 0.15$, $h^\female = h^\male = 0.7$, $t^\female = 0$, $t^\male = -0.1$, $\alpha^\male = \alpha^\female = 1/2$, $r=0.01$, $R=0.05$.
%}
%\label{fig:WinvasionLowR}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%W invades XY but does not fix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{FreqWHighR}\\
%\includegraphics[width=0.5\linewidth]{SexRatioHighR}\\
%\includegraphics[width=0.5\linewidth]{MeanDipFitHighR}
%\caption{
%Haploid selection allows a completely unlinked neo-$W$ to invade an ancestral $XY$ system (\textbf{A}) despite further biasing the sex ratio (\textbf{B}) and decreasing mean diploid viability fitness (\textbf{C}).
%The neo-$W$ does not fix (although variation at the \textbf{A} locus is maintained, $V_A>0$), resulting in a polymorphic sex-determination system.
%Parameters as in Figure \ref{fig:WinvasionLowR} but with $R=0.5$.
%}
%\label{fig:WinvasionHighR}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%W invades for all rates of recombination with selected locus
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{InvasionVsRecombination}
%\caption{
%A neo-$W$ invades an ancestral $XY$ system with haploid selection regardless of how tightly it is linked to the selected locus. 
%Parameters as in Figure \ref{fig:WinvasionLowR}.
%}
%\label{fig:InvasionVsRecombination}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%W invasion versus position of selected locus
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{InvasionVsCentiMorgans}
%\includegraphics[width=0.5\linewidth]{InvasionVsCentiMorgansAllOrders}
%\caption{
%Haploid selection allows a neo-$W$ to invade an ancestral $XY$ system regardless of how tightly it and the ancestral sex-determining locus are linked to the selected locus. 
%The ancestral sex-determining locus is located at -0.05 and the novel sex-determining locus is located at 0.05 (corresponding to the peaks of invasion fitness), such that the probability of a cross-over between them is $\approx0.1$.
%The x-axis gives the position of the locus under haploid selection.
%We used Haldane's map function \citep[Equation 3 in ][]{Haldane1919} to convert from map distance (centiMorgans) to the probability of a cross-over event. 
%Parameters as in Figure \ref{fig:WinvasionLowR}.
%}
%\label{fig:InvasionVsCentiMorgans}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Y invades ZW and fixes (counter example to Kozi explanation)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}
%\centering
%\includegraphics[width=0.5\linewidth]{FreqWCounterKozi}\\
%\includegraphics[width=0.5\linewidth]{SexRatioCounterKozi}\\
%\includegraphics[width=0.5\linewidth]{MeanDipFitCounterKozi}
%\caption{
%Meiotic drive allows a neo-$Y$ to invade an ancestral $ZW$ system and fix (\textbf{A}) despite temporarily biasing the sex ratio (\textbf{B}). %and decreasing mean diploid viability fitness (\textbf{C}).
%Parameters: $k=0$, $s^\female = s^\male = t^\female = t^\male = 0$, $\alpha^\male = 0.4$, $\alpha^\female = 1/2$, $r=0$, $R=0$.
%}
%\label{fig:CounterKozi}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\setcounter{equation}{0}
\renewcommand{\theequation}{S.\arabic{equation}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{S.\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{S.\arabic{table}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage

\section*{Appendix}

\subsection*{Recursion Equations}
\label{app:recurs}

\textcolor{red}{Should we adjust the subscripts throughout this subsection? Right now we end up re-defining $i$ and $j$ (when switching from haploid to diploid; this might have been my doing!) and then introduce three new subscripts $b$, $c$, and $l$, all of which can be derived from $i$ and $j$. Might be more straightforward to just use $p^\Hermaphrodite_{x_1,x_2,a_1,a_2,m_1,m_2}$ where 1 is maternal and 2 is paternal? We then no longer have to switch indices from haploid to diploid and the connection to other variables is clear: $b=m_1m_2$, $c=x_1x_2$, and $l=a_1a_2$. I guess the downside will be re-writing the recursion equations... which is why I haven't gone ahead and tried this.}

In each generation we census the genotype frequencies in male and female gametes/gametophytes (hereafter, gametes) between meiosis (and any meiotic drive) and gametic competition. 
At this stage we denote the frequencies of X- and Y-bearing gametes from males and females $x_{ij}^{\Hermaphrodite}$ and $y_{ij}^{\Hermaphrodite}$, where $\Hermaphrodite \in \{\male,\female\}$ specifies the sex of the diploid that the gamete came from, $i\in\{A,a\}$ specifies the allele at the selected locus $\mathbf{A}$, and $j\in\{M,m\}$ specifies the allele at the novel sex-determining locus $\mathbf{M}$. 
The gamete frequencies from each sex sum to one, $\sum_{i,j}x_{ij}^{\Hermaphrodite}+y_{ij}^{\Hermaphrodite}=1$.

Competition then occurs among gametes of the same sex (e.g., among eggs and among sperm separately) according to the \textbf{A} locus allele, $i$ (see Table \ref{tab:fitnesstable}).
The genotype frequencies after gametic competition are $x_{ij}^{\Hermaphrodite,s}= w_{i}x_{ij}^{\Hermaphrodite}/\bar{w}_{H}^{\Hermaphrodite}$ and $y_{ij}^{\Hermaphrodite,s}= w_{i}y_{ij}^{\Hermaphrodite}/\bar{w}_{H}^{\Hermaphrodite}$, where $\bar{w}_{H}^{\Hermaphrodite}=\sum_{i,j} w_{i}x_{ij}^{\Hermaphrodite}+w_{i}y_{ij}^{\Hermaphrodite}$ is the mean fitness of male ($\Hermaphrodite=\male$) or female ($\Hermaphrodite=\female$) gametes.

Random mating then occurs between gametes to produce diploid zygotes.
To shorten notation we now use index $i$ (and $j$) to denote the alleles at both the $\mathbf{A}$ and $\mathbf{M}$ loci and label $MA=1$, $Ma=2$, $mA=3$, and $ma=4$, such that $i,j\in\{1,2,3,4\}$.
The frequencies of $XX$ zygotes are then denoted as $xx_{ij}$, $XY$ zygotes as $xy_{ij}$, and $YY$ zygotes as $yy_{ij}$. 
In $XX$ and $YY$ zygotes, individuals with diploid genotype $ij$ are equivalent to those with diploid genotype $ji$; for simplicity, we use $xx_{ij}$ and $yy_{ij}$ with $i\neq j$ to denote the average of these frequencies, $xx_{ij}=(x_{i}^{\female,s}x_{j}^{\male,s}+x_{j}^{\female,s}x_{i}^{\male,s})/2$ and $yy_{ij}=(y_{i}^{\female,s}y_{j}^{\male,s}+y_{j}^{\female,s}y_{i}^{\male,s})/2$.

Denoting the \textbf{M} locus genotype by $b \in \{MM, Mm, mm\}$ and the \textbf{X} locus genotype by $c \in \{XX, XY, YY\}$, zygotes develop as females with probability $k_{bc}$. 
Therefore, the frequencies of $XX$ females are given by $xx_{ij}^{\female}=k_{bc}xx_{ij}$, $XY$ females are given by $xy_{ij}^{\female}=k_{bc}xy_{ij}$, and $YY$ females are given by $yy_{ij}^{\female}=k_{bc}yy_{ij}$. 
Similarly, $XX$ male frequencies are $xx_{ij}^{\male}=(1-k_{bc})xx_{ij}$, $XY$ male frequencies are $xy_{ij}^{\male}=(1-k_{bc})xy_{ij}$, and $YY$ males frequencies are $yy_{ij}^{\male}=(1-k_{bc})yy_{ij}$.
This notation allows both the ancestral and novel sex-determining regions to determine zygotic sex according to an $XY$ system, a $ZW$ system, or an environmental sex-determining system. 
In addition, we can consider any epistatic dominance relationship between the two sex-determining loci. 
For example, here we assume that the ancestral sex-determining system (\textbf{X} locus) is $XY$ ($k_{MMXX}=1$ and $k_{MMXY}=k_{MMYY}=0$) or $ZW$ ($k_{MMZZ}=0$ and $k_{MMZW}=k_{MMWW}=1$) and epistatically recessive to a dominant novel sex-determining locus, \textbf{M} ($k_{Mmc}=k_{mmc}=k$).

Selection among diploids then occurs according to the diploid genotype at the \textbf{A} locus, $l \in \{AA, Aa, aa\}$, for an individual of type $ij$ (see Table \ref{tab:fitnesstable}). 
The diploid frequencies after selection in sex $\Hermaphrodite$ are given by $xx_{ij}^{\Hermaphrodite,s}=w_{l}^{\Hermaphrodite} xx_{ij}/\bar{w}^{\Hermaphrodite}$, $xy_{ij}^{\Hermaphrodite,s}=w_{l}^{\Hermaphrodite} xy_{ij}/\bar{w}^{\Hermaphrodite}$, and $yy_{ij}^{\Hermaphrodite,s}=w_{l}^{\Hermaphrodite} yy_{ij}/\bar{w}^{\Hermaphrodite}$, where $\bar{w}^{\Hermaphrodite}= \sum_{i=1}^{4}\sum_{j=1}^{4}w_{l}^{\Hermaphrodite}xx_{ij}+w_{l}^{\Hermaphrodite}xy_{ij}+w_{l}^{\Hermaphrodite}yy_{ij}$ is the mean fitness of individuals of sex $\Hermaphrodite$.

Finally, these diploids undergo meiosis to produce the next generation of gametes. 
Recombination and sex-specific meiotic drive occur during meiosis.
%We can also assume that recombination is sex specific and/or affected by the M locus - but generally we don't so I just describe $R=R_{f}=R, r_{MM,d}=r_{Mm,d}=r_{mm,d}=r, \chi_{m}=\chi_{f}=\chi$.
Here, we allow any relative locations for the SDR, \textbf{A}, and \textbf{M} loci by using three parameters to describe the recombination rates between them. 
$R$ is the recombination rate between the \textbf{A} locus and the \textbf{M} locus, $\chi$ is the recombination rate between the \textbf{M} locus and the \textbf{X} locus, and $r$ is the recombination rate between the \textbf{A} locus and the \textbf{X} locus. 
Table \ref{tab:chisubstitutions} shows how $\chi$ can be substituted to give any linear order of loci.
During meiosis in sex $\Hermaphrodite$, meiotic drive occurs such that, in $Aa$ heterozygotes, a fraction $\alpha^{\Hermaphrodite}$ of gametes produced carry the $A$ allele and $(1-\alpha^\Hermaphrodite)$ carry the $a$ allele.

\begin{table}[ht]
\centering
\smallskip
\caption{$\chi$ substitutions for different loci orders (assuming no interference)}
\begin{tabular}{l l}
\hline\hline
  Order of loci &   \\ [0.5ex] \hline
  SDR-A-M & $\chi=R(1-r)+r(1-R)$  \\
  SDR-M-A & $\chi=(r-R)/(1-2R)$ \\
  A-SDR-M & $\chi=(R-r)/(1-2r)$ \\
  \hline \hline
  \label{tab:chisubstitutions}
 \end{tabular}
\end{table}

Among gametes from sex $\Hermaphrodite$ (sperm/pollen when $\Hermaphrodite=\male$, eggs/ovules when $\Hermaphrodite=\female$), the frequencies of haplotypes (before gametic competition) in the next generation are given by

\begingroup
\allowdisplaybreaks
\begin{subequations}
\begin{align}
%
\begin{split}
x_{MA}^{{\Hermaphrodite}'}=&xx_{11}^{\Hermaphrodite,s}+xx_{13}^{\Hermaphrodite,s}/2+(xx_{12}^{\Hermaphrodite,s}+xx_{14}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(xx_{14}^{\Hermaphrodite,s}-xx_{23}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&+(xy_{11}^{\Hermaphrodite,s}+xy_{13}^{\Hermaphrodite,s})/2+(xy_{12}^{\Hermaphrodite,s}+xy_{14}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&- r(xy_{12}^{\Hermaphrodite,s}-xy_{21}^{\Hermaphrodite,s})\alpha^\Hermaphrodite - \chi(xy_{13}^{\Hermaphrodite,s}-xy_{31}^{\Hermaphrodite,s})/2\\
&+\big{\{}-(R+r+\chi)xy_{14}^{\Hermaphrodite,s} +(r+\chi-R)xy_{41}^{\Hermaphrodite,s}\\
&+(R+r-\chi) xy_{23}^{\Hermaphrodite,s}+(R+\chi-r)xy_{32}^{\Hermaphrodite,s}\big{\}}\alpha^\Hermaphrodite/2
\end{split}
\\
%
\begin{split}
x_{Ma}^{{\Hermaphrodite}'}=&xx_{22}^{\Hermaphrodite,s}+xx_{24}^{\Hermaphrodite,s}/2+(xx_{12}^{\Hermaphrodite,s}+xx_{23}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(xx_{23}^{\Hermaphrodite,s}-xx_{14}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{22}^{\Hermaphrodite,s}+xy_{24}^{\Hermaphrodite,s})/2+(xy_{21}^{\Hermaphrodite,s}+xy_{23}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite)\\
&- r(xy_{21}^{\Hermaphrodite,s}-xy_{12}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite) - \chi(xy_{24}^{\Hermaphrodite,s}-xy_{42}^{\Hermaphrodite,s})/2\\
&+\big{\{}-(R+r+\chi)xy_{23}^{\Hermaphrodite,s}+(r+\chi-R)xy_{32}^{\Hermaphrodite,s}\\
&+(R+r-\chi) xy_{14}^{\Hermaphrodite,s}+(R+\chi-r)xy_{41}^{\Hermaphrodite,s}\big{\}}(1-\alpha^\Hermaphrodite)/2
\end{split}
\\
%
\begin{split}
x_{mA}^{{\Hermaphrodite}'}=&xx_{33}^{\Hermaphrodite,s}+xx_{13}^{\Hermaphrodite,s}/2+(xx_{23}^{\Hermaphrodite,s}+xx_{34}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(xx_{23}^{\Hermaphrodite,s}-xx_{14}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{33}^{\Hermaphrodite,s}+xy_{31}^{\Hermaphrodite,s})/2+(xy_{32}^{\Hermaphrodite,s}+xy_{34}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&- r(xy_{34}^{\Hermaphrodite,s}-xy_{43}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite - \chi(xy_{31}^{\Hermaphrodite,s}-xy_{13}^{\Hermaphrodite,s})/2\\
&+\big{\{}-(R+r+\chi)xy_{32}^{\Hermaphrodite,s} +(r+\chi-R)xy_{23}^{\Hermaphrodite,s}\\
&+(R+r-\chi) xy_{41}^{\Hermaphrodite,s}+(R+\chi-r)xy_{14}^{\Hermaphrodite,s}\big{\}} \alpha^\Hermaphrodite/2
\end{split}
\\
%
\begin{split}
x_{ma}^{{\Hermaphrodite}'}=&xx_{44}^{\Hermaphrodite,s}+xx_{34}^{\Hermaphrodite,s}/2+(xx_{14}^{\Hermaphrodite,s}+xx_{24}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(xx_{14}^{\Hermaphrodite,s}-xx_{23}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{44}^{\Hermaphrodite,s}+xy_{42}^{\Hermaphrodite,s})/2+(xy_{41}^{\Hermaphrodite,s}+xy_{43}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite)\\
&- r(xy_{43}^{\Hermaphrodite,s}-xy_{34}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite) - \chi(xy_{42}^{\Hermaphrodite,s}-xy_{24}^{\Hermaphrodite,s})/2\\
&+\big{\{}-(R+r+\chi)xy_{41}^{\Hermaphrodite,s}+(r+\chi-R)xy_{14}^{\Hermaphrodite,s}\\
&+(R+r-\chi) xy_{32}^{\Hermaphrodite,s}+(R+\chi-r)xy_{23}^{\Hermaphrodite,s}\big{\}}(1-\alpha^\Hermaphrodite)/2
\end{split}
\\
\begin{split}
y_{MA}^{{\Hermaphrodite}'}=&yy_{11}^{\Hermaphrodite,s}+yy_{13}^{\Hermaphrodite,s}/2+(yy_{12}^{\Hermaphrodite,s}+yy_{14}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(yy_{14}^{\Hermaphrodite,s}-yy_{23}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{11}^{\Hermaphrodite,s}+xy_{31}^{\Hermaphrodite,s})/2+(xy_{21}^{\Hermaphrodite,s}+xy_{41}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&- r(xy_{21}^{\Hermaphrodite,s}-xy_{12}^{\Hermaphrodite,s})\alpha^\Hermaphrodite - \chi(xy_{31}^{\Hermaphrodite,s}-xy_{13}^{\Hermaphrodite,s})/2\\
&+\big{\{} - (R+r+\chi)xy_{41}^{\Hermaphrodite,s} + (r+\chi-R)xy_{14}^{\Hermaphrodite,s}\\
& + (R+r-\chi)xy_{32}^{\Hermaphrodite,s} + (R+\chi-r) xy_{23}^{\Hermaphrodite,s}\big{\}}\alpha^\Hermaphrodite/2
\end{split}
\\
%
\begin{split}
y_{Ma}^{{\Hermaphrodite}'}=&yy_{22}^{\Hermaphrodite,s}+yy_{24}^{\Hermaphrodite,s}/2+(yy_{12}^{\Hermaphrodite,s}+yy_{23}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(yy_{23}^{\Hermaphrodite,s}-yy_{14}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{22}^{\Hermaphrodite,s}+xy_{42}^{\Hermaphrodite,s})/2+(xy_{12}^{\Hermaphrodite,s}+xy_{32}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite)\\
&- r(xy_{12}^{\Hermaphrodite,s}-xy_{21}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite) - \chi(xy_{42}^{\Hermaphrodite,s}-xy_{24}^{\Hermaphrodite,s})/2\\
&+\big{\{}-(R+r+\chi)xy_{32}^{\Hermaphrodite,s} +(r+\chi-R) xy_{23}^{\Hermaphrodite,s}\\
&+(R+r-\chi)xy_{41}^{\Hermaphrodite,s}+(R+\chi-r)xy_{14}^{\Hermaphrodite,s}\big{\}}(1-\alpha^\Hermaphrodite)/2
\end{split}
\\
%
\begin{split}
y_{mA}^{{\Hermaphrodite}'}=&yy_{33}^{\Hermaphrodite,s}+yy_{13}^{\Hermaphrodite,s}/2+(yy_{23}^{\Hermaphrodite,s}+yy_{34}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(yy_{23}^{\Hermaphrodite,s}-yy_{14}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{33}^{\Hermaphrodite,s}+xy_{13}^{\Hermaphrodite,s})/2+(xy_{23}^{\Hermaphrodite,s}+xy_{43}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&- r(xy_{43}^{\Hermaphrodite,s}-xy_{34}^{\Hermaphrodite,s})\alpha^\Hermaphrodite - \chi(xy_{13}^{\Hermaphrodite,s}-xy_{31}^{\Hermaphrodite,s})/2\\
&+\big{\{}-(R+r+\chi)xy_{23}^{\Hermaphrodite,s} +(r+\chi-R)xy_{32}^{\Hermaphrodite,s}\\
&+(R+r-\chi) xy_{14}^{\Hermaphrodite,s} + (R+\chi-r) xy_{41}^{\Hermaphrodite,s}\big{\}}\alpha^\Hermaphrodite/2
\end{split}
\\
%
\begin{split}
y_{ma}^{{\Hermaphrodite}'}=&yy_{44}^{\Hermaphrodite,s}+yy_{34}^{\Hermaphrodite,s}/2+(yy_{14}^{\Hermaphrodite,s}+yy_{24}^{\Hermaphrodite,s})\alpha^\Hermaphrodite\\
&-R(yy_{14}^{\Hermaphrodite,s}-yy_{23}^{\Hermaphrodite,s}) \alpha^\Hermaphrodite\\
&(xy_{44}^{\Hermaphrodite,s}+xy_{24}^{\Hermaphrodite,s})/2+(xy_{14}^{\Hermaphrodite,s}+xy_{34}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite)\\
&- r(xy_{34}^{\Hermaphrodite,s}-xy_{43}^{\Hermaphrodite,s})(1-\alpha^\Hermaphrodite) - \chi(xy_{24}^{\Hermaphrodite,s}-xy_{42}^{\Hermaphrodite,s})/2\\
&+\big{\{}-(R+r+\chi) xy_{14}^{\Hermaphrodite,s} + (r+\chi-R)xy_{41}^{\Hermaphrodite,s}\\
&+(R+r-\chi) xy_{23}^{\Hermaphrodite,s} + (R+\chi-r) xy_{32}^{\Hermaphrodite,s}\big{\}}(1-\alpha^\Hermaphrodite)/2
\end{split}
\end{align}
\label{eq:recursions}
\end{subequations}

\endgroup

\noindent
The full system is therefore described by 16 recurrence equations (three diallelic loci in two sexes, $2^3 \times 2 = 16$). 
However, some diploid types are not produced under a given sex-determination system. 
For example, with the $M$ allele fixed and ancestral $XY$ sex determination, there are no $m$ alleles, $XX$ males, $XY$ females, or $YY$ females ($xx_{11}^{\male}=xx_{12}^{\male}=xx_{22}^\male=xy_{11}^{\female}=xy_{12}^{\female}=xy_{21}^{\female}=xy_{22}^\female=yy_{11}^{\female}=yy_{12}^{\female}=yy_{22}^\female=0$). 
In this case, the system only involves six recursion equations, %because there is only one \textbf{M} locus allele and no Y-bearing female gametes. 
which yields equilibrium \eqref{eq:pAve}.

%I think this should be moved below because we have some 'results' where we don't assume weak selection but we only have the equilibrium calculated analytically for weak selection.

\subsection*{Resident equilibrium and stability}

In the resident population (allele $M$ fixed), we choose to follow the frequency of $A$ in female gametes (eggs) from an XX female, $p^\female_X$, and in X-bearing, $p^\male_X$, and Y-bearing, $p^\male_Y$, male gametes (sperm).
We also track the total frequency of Y among male gametes, $q$, which may deviate from $1/2$ due to meiotic drive in males. 
These four variables determine the frequencies of the six resident gamete types: $x_{MA}^{\female}=p_{X}^{\female}$, $x_{Ma}^{\female}=1-p_{X}^{\female}$, $x_{MA}^{\male}=(1-q)p_{X}^{\male}$, $x_{Ma}^{\male}=(1-q)(1-p_{X}^{\male})$, $y_{MA}^{\male}=q p_{Y}^{\male}$, and $y_{Ma}^{\male}=q(1-p_{Y}^{\male})$. 
Mean fitnesses in the resident population are given in table \ref{tab:meanfitnesses}.

Various forms of selection can maintain a polymorphism at the \textbf{A} locus, including sexually antagonistic selection, overdominance, conflicts between diploid selection and selection upon haploid genotypes \citep[ploidally antagonistic selection,][]{Immler:2012tl}, and a combination of these selective regimes.

\begin{table}[ht]
\centering
\smallskip
\caption{mean fitnesses in the resident population ($M$ fixed, XY sex determination) }
\begin{tabular}{l l }
\hline\hline
  Sex \& Life Cycle Stage & Mean Fitness \\ [0.5ex] \hline  \noalign{\vskip 0.5ex}
  female gametes ($\bar{w}_H^\female$) & 
  $p_X^\female w_A^\female + (1-p_X^\female) w_a^\female$ \\ [0.5ex] \hline  \noalign{\vskip 0.5ex}
  male gametes ($\bar{w}_H^\male$) & 
  $\bar{p}^{\male} w_A^\male + (1-\bar{p}^{\male}) w_a^\male$ \\ [0.5ex] \hline  \noalign{\vskip 0.5ex}
  females ($\bar{w}^\female$) & 
  $\begin{array}{l}  \{ p_X^\female w_A^\female p_X^\male w_A^\male w_{AA}^\female + \\
  (1 - p_X^\female) w_a^\female p_X^\male w_A^\male w_{Aa}^\female + \\
  p_X^\female w_A^\female (1 - p_X^\male) w_a^\male w_{Aa}^\female + \\
  (1-p_X^\female) w_a^\female (1 - p_X^\male) w_a^\male w_{aa}^\female \} / \{ \bar{w}_H^\female \bar{w}_H^\male \zeta \}
  \end{array} 
  $ \\ [0.5ex] \hline  \noalign{\vskip 0.5ex}
  males ($\bar{w}^\male$) & 
  $\begin{array}{l} \{ p_X^\female w_A^\female  p_Y^\male w_A^\male w_{AA}^\male + \\
  (1 - p_X^\female) w_a^\female  p_Y^\male w_A^\male w_{Aa}^\male + \\
  p_X^\female w_A^\female  (1 - p_Y^\male) w_a^\male w_{Aa}^\male + \\
  (1-p_X^\female) w_a^\female  (1 - p_Y^\male) w_a^\male w_{aa}^\male \} / \{ \bar{w}_H^\female \bar{w}_H^\male (1-\zeta)\} 
  \end{array}
  $ \\ [0.5ex] \hline  \noalign{\vskip 0.5ex}
  zygotic sex ratio $\zeta$ & $\{ (1-q)(p_X^\male w_A^\male+(1-p_X^\male)w_a^\male)\} / \bar{w}_H^\male $
   \\ [0.5ex]  \noalign{\vskip 0.5ex}
  \hline \hline
  \label{tab:meanfitnesses}
 \end{tabular}
\end{table}

In particular special cases, e.g., no sex-differences in selection or meiotic drive ($s^\male=s^\female$, $h^\male=h^\female$, and $\alpha^\male=\alpha^\female=1/2$), the equilibrium allele frequency and stability can be calculated analytically without assuming anything about the relative strengths of selection and recombination. 
However, here, we focus on two regimes (tight linkage and weak selection) in order to make fewer assumptions about fitnesses.

\subsubsection*{Recombination weak relative to selection (tight linkage between \textbf{A} and \textbf{X})}

We first calculate the equilibrium frequency of the Y and $A$ alleles in the ancestral population when the recombination rate between the \textbf{X} and \textbf{A} loci is small ($r$ of order $\epsilon$). 
The \textbf{A} locus will not affect evolution at the novel sex-determining locus, \textbf{M}, if one \textbf{A} allele is fixed on all backgrounds. 
We therefore focus on the five equilibria that maintain both $A$ and $a$ alleles, four of which are given to leading order by:

\begin{subequations}\label{eq:tightequil}
\begin{align}
(A)\ \ \ &\hat{p}_{Y}^{\male}=0,
\ \hat{q}=\frac{1}{2}-\frac{(\alpha^{\male}-1/2)w_{Aa}^{\male} \phi}{w_{Aa}^{\male} \phi+ w_{aa}^{\male} \psi},\\
&\hat{p}_{X}^{\female}=\frac{w_{a}^{\female} \phi}{w_{a}^{\female} \phi+ w_{A}^{\female} \psi},
\ \hat{p}_{X}^{\male}=\frac{2 \alpha^{\male}w_{Aa}^{\male} \phi}{2\alpha^{\male}w_{Aa}^{\male} \phi +w_{AA}^{\male} \psi} \nonumber \\
(A')\ \ \ &\hat{p}_{Y}^{\male}=1,
\ \hat{q}=\frac{1}{2}+\frac{(\alpha^{\male}-1/2)w_{Aa}^{\male} \phi'}{w_{Aa}^{\male} \phi' + w_{AA}^{\male} \psi'},\\
&\hat{p}_{X}^{\female}=1-\frac{w_{A}^{\female} \phi'}{w_{A}^{\female} \phi'+ w_{a}^{\female} \psi'},
\ \hat{p}_{X}^{\male}=1-\frac{2(1-\alpha^{\male})w_{Aa}^{\male} \phi'}{2(1-\alpha^{\male})w_{Aa}^{\male} \phi'+w_{aa}^{\male} \psi'} \nonumber \\
(B)\ \ \ &\hat{p}_{Y}^{\male}=0,\ \hat{p}_{X}^{\female}=1,\ \hat{p}_{X}^{\male}=1, \ \hat{q}=1-\alpha^{\male}\\
(B')\ \ \ &\hat{p}_{Y}^{\male}=1,\ \hat{p}_{X}^{\female}=0,\ \hat{p}_{X}^{\male}=0, \ \hat{q}=\alpha^{\male}
\end{align}
\end{subequations}
\begin{equation*}
\begin{split}
\phi=&\alpha^{\female} w_{A}^{\female} w_{Aa}^{\female}(w_{a}^{\male} w_{aa}^{\male} + 2 \alpha^{\male} w_{A}^{\male} w_{Aa}^{\male}) - w_{a}^{\male} w_{a}^{\female} w_{aa}^{\male} w_{aa}^{\female} \\
\psi=&(1-\alpha^{\female}) w_{a}^{\female} w_{Aa}^{\female}(w_{a}^{\male} w_{aa}^{\male} + 2 \alpha^{\male} w_{A}^{\male} w_{Aa}^{\male}) - 2\alpha^{\male} w_{A}^{\male} w_{A}^{\female} w_{Aa}^{\male} w_{AA}^{\female}\\
\phi'=&(1-\alpha^{\female}) w_{a}^{\female} w_{Aa}^{\female}(w_{A}^{\male} w_{AA}^{\male} + 2 (1-\alpha^{\male}) w_{a}^{\male} w_{Aa}^{\male}) - w_{A}^{\male} w_{A}^{\female} w_{AA}^{\male} w_{AA}^{\female}\\
\psi'=&\alpha^{\female} w_{A}^{\female} w_{Aa}^{\female}(w_{A}^{\male} w_{AA}^{\male} + 2 (1-\alpha^{\male}) w_{a}^{\male} w_{Aa}^{\male}) - 2(1-\alpha^{\male}) w_{a}^{\male} w_{a}^{\female} w_{Aa}^{\male} w_{aa}^{\female}
\end{split}
\end{equation*}

\noindent
A fifth equilibrium $(C)$ also exists where $A$ is present at an intermediate frequency on the Y chromosome ($0<\hat{p}_{Y}^{\male}<1$). However, equilibrium $(C)$ is never locally stable when $r \approx 0$ and is therefore not considered further. 
Thus, the Y can either be fixed for the $a$ allele (equilibria $A$ and $B$) or the $A$ allele (equilibria $A'$ and $B'$).
The X chromosome can then either be polymorphic (equilibria $A$ and $A'$) or fixed for the alternative allele (equilibria $B$ and $B'$).
Since equilibria $(A)$ and $(B)$ are equivalent to equilibria $(A')$ and $(B')$ with the labelling of $A$ and $a$ alleles interchanged, we discuss only equilibria $(A)$ and $(B)$, in which the Y is fixed for the $a$ allele. 
If there is no haploid selection ($\alpha^{\Hermaphrodite}=1/2$, $w_{A}^{\Hermaphrodite}=w_{a}^{\Hermaphrodite}=1$), these equilibria are equivalent to those found by \cite{Lloyd1977} and \cite{Otto2014}.

We next calculate when $(A)$ and $(B)$ are locally stable for $r=0$. According to the `small parameter theory' \citep{Karlin:1972ab,Karlin:1972dq}, these stability properties are unaffected by small amounts of recombination between the SDR and \textbf{A} locus, although equilibrium frequencies may be slightly altered. 
For the $a$ allele to be stably fixed on the Y requires that $\bar{w}_{Ya}^{\male}>\bar{w}_{YA}^{\male}$ where $\bar{w}_{Ya}^{\male}=w_{a}^{\male}(2 p_{X}^{\female}(1-\alpha^{\male}) w_{A}^{\female} w_{Aa}^{\male}+ (1-p_{X}^{\female})w_{a}^{\female} w_{aa}^{\male} )$ and $\bar{w}_{YA}^{\male}=w_{A}^{\male}(p_{X}^{\female} w_{A}^{\female} w_{AA}^{\male}+ 2 (1-p_{X}^{\female})\alpha^{\male}w_{a}^{\female} w_{Aa}^{\male} )$. That is, $Ya$ haplotypes must have higher fitness than $YA$ haplotypes.  
Substituting in $p_{X}^{\female} = \hat{p}_{X}^{\female}$ from above, fixation of the $a$ allele on the Y requires that $\gamma_{i}>0$ where $\gamma_{(A)}=w_{a}^{\male}(2(1-\alpha^{\male})w_{Aa}^{\male} \phi + w_{aa}^{\male} \psi)-w_{A}^{\male}( w_{AA}^{\male} \phi + 2\alpha^{\male}w_{Aa}^{\male} \psi)$ for equilibrium $(A)$ and $\gamma_{(B)}=2(1-\alpha^{\male})w_{a}^{\male}w_{Aa}^{\male}-w_{A}^{\male}w_{AA}^{\male}$ for equilibrium $(B)$.
Stability of a polymorphism on the X chromosome (equilibrium $A$) further requires that $\phi >0$ and $\psi >0$. 
Fixation of the $a$ allele on the X (equilibrium $B$) is mutually exclusive with equilibrium $(A)$ and requires $\psi<0$ and $w_{A}^{\female}w_{AA}^{\female}>(1-\alpha^{\female})w_{a}^{\female}w_{Aa}^{\female}$.

\subsubsection*{Selection weak relative to recombination (weak selection)}

Here, we assume that selection and meiotic drive are weak relative to recombination ($s^\Hermaphrodite$, $t^\Hermaphrodite$, $\alpha_{\Delta}^\Hermaphrodite$ of order $\epsilon$). 
The maintenance of a polymorphism at the \textbf{A} locus then requires that

\begin{equation}
\begin{split}
0&<-((1-h^\female)s^\female +(1-h^\male) s^\male + t^\female +t^\male + \alpha_{\Delta}^\female+\alpha_{\Delta}^\male)\\
%
\text{and }\quad 0&<(h^\female s^\female +h^\male s^\male + t^\female +t^\male + \alpha_{\Delta}^\female+\alpha_{\Delta}^\male).
\end{split}
\end{equation}

\noindent
which indicates that a polymorphism can be maintained by various selective regimes.

Given that a polymorphism is maintained at the \textbf{A} locus by selection, with weak selection and drive the frequencies of $A$ in each type of gamete are the same ($\hat{p}^\female_X=\hat{p}^\male_X=\hat{p}^\male_Y=\bar{p}$) and given, to leading order, by

\begin{equation}
\bar{p}=\frac{h^\female s^\female + h^\male s^\male +t^\female+t^\male+\alpha_{\Delta}^\female+\alpha_{\Delta}^\male}
{(2h^\female-1)s^\female+(2h^\male-1)s^\male}
+O(\epsilon)
.
\label{eq:pAve}
\end{equation}

\noindent
Differences in frequency between gamete types are of order $\epsilon$ and given, to leading order, by

\begin{equation}
\begin{split}
\hat{p}^\male_X-\hat{p}^\female_X&=V_{A}\big{(}D^\male - D^\female +\alpha_{\Delta}^\male-\alpha_{\Delta}^\female \big{)}
+O(\epsilon^2)\\
%
\hat{p}^\male_Y-\hat{p}^\female_X&=V_{A}\big{(}D^\male - D^\female+\alpha_{\Delta}^\male-\alpha_{\Delta}^\female+(1-2r)(t^\male-t^\female)\big{)}/2r
+O(\epsilon^2)\\
%
\hat{p}^\male_Y-\hat{p}^\male_X&=V_{A}\big{(}D^\male - D^\female+\alpha_{\Delta}^\male-\alpha_{\Delta}^\female+t^\male-t^\female\big{)}(1-2r)/2r
+O(\epsilon^2)
\end{split}
\label{eq:freq_diffs}
\end{equation}

\noindent
where $V_{A}=\bar{p}(1-\bar{p})$ is the variance in the frequency of $A$ and $D^\Hermaphrodite=\big{(} \bar{p}s^\Hermaphrodite+(1-\bar{p})h^\Hermaphrodite s^\Hermaphrodite\big{)} -\big{(} \bar{p}h^\Hermaphrodite s^\Hermaphrodite+(1-\bar{p}) \big{)}$ corresponds to the difference in fitness between $A$ and $a$ alleles in diploids of sex $\Hermaphrodite \in \{\female,\male\}$ ($\bar{p}$ is the leading-order probability of mating with an $A$-bearing gamete from the opposite sex). 
The frequency of Y among male gametes depends upon the difference in the frequency of the $A$ allele between X- and Y-bearing male gametes and the strength of meiotic drive in favour of the $A$ allele in males, $q=1/2+\alpha_{\Delta}^\male(\hat{p}^\male_Y-\hat{p}^\male_X)/2+O(\epsilon^3)$.
Without gametic competition or drive ($\alpha_{\Delta}^\Hermaphrodite=t^\Hermaphrodite=0$) our results reduce to those of \citet{vanDoorn:2007eu}.

% the differences in the frequencies of $A$ in each type of gamete are small, as is the bias in the sex-determining factor from the heterogametic sex, and we can solve for the mean frequency of $A$ across all types ($p_A$), the difference in the frequencies of $A$ between two of the three types, and the bias in the frequency of the sex-determining factor, to first order in selection.
%Linear stability analysis can then be used to determine the stability of this equilibrium.
%Without haploid selection or meiotic drive our results reduce to those of \cite{vanDoorn:2007eu}. %when $k=0$ (neo-$Y$ invading a $ZW$ system) or when $k=1$ (neo-$W$ invading an $XY$ system) (\textcolor{red}{CHECK}).
%However, with haploid selection or meiotic drive a stable polymorphism at locus \textbf{A} no longer requires sexually antagonistic selection. %, as it can also be achieved by ploidally-antagonistic selection.

\subsection*{Invasion conditions}

%Here, we determine whether a rare neo-Y or neo-W allele spreads when rare, which occurs when $\lambda > 1$.
%We begin with the general result and then give explicit solutions under tight linkage and weak selection.

%If the average change in frequency of the two haplotypes that carry the $m$ allele ($Am$ and $am$) is positive, invasion will always occur (i.e., if $\left\{(\lambda_{mA}-1)+ (\lambda_{ma}-1) \right\}/2 > 0$ then $\lambda > 1$, see table \ref{tab:haplotype_growth} for $\lambda_{mi}$). 
%If neither haplotype increases in frequency ($\lambda_{mA}, \lambda_{ma} < 1$), the $m$ allele will not invade. 
%Otherwise, the new sex-determining allele increases in frequency on one \textbf{A} background and declines on the other, and invasion requires
%\begin{equation}\label{eq:neoYR}
%R \left[ 
%\frac{p^\female_X w^\female_A w^\male_a (1-\alpha^\male)}{ \bar{w}^\female_H \bar{w}^\male_H (\lambda_{mA}-1)} + 
%p^\female_X \frac{w^\female_A}{\bar{w}^\female_H} \frac{w^\male_a(1-\alpha^\male)}{\bar{w}^\male_H} \frac{1}{\lambda_{mA}-1} + 
%\frac{(1 - p^\female_X) w^\female_a w^\male_A \alpha^\male}{ \bar{w}^\female_H \bar{w}^\male_H (\lambda_{ma}-1)} 
%\right] 
%\frac{w^\male_{Aa}}{q\bar{w}^\male} < 1,
%\end{equation}

%\noindent
%for the neo-$Y$, and

%\noindent
%\begin{equation}\label{eq:neoWR}
%R \left[ 
%\frac{\bar{p}^\male w^\male_A w^\female_a (1-\alpha^\female)}{ \bar{w}^\male_H \bar{w}^\female_H (\lambda_{mA}-1)} + 
%\frac{(1 - \bar{p}^\male) w^\male_a w^\female_A \alpha^\female}{ \bar{w}^\male_H \bar{w}^\female_H (\lambda_{ma}-1)} 
%\right] 
%\frac{w^\female_{Aa}}{(1-q)\bar{w}^\female} < 1,
%\end{equation}

%\noindent
%for the neo-$W$.

%Equations \eqref{eq:neoYR} and \eqref{eq:neoWR} show that the new sex-determining allele, $m$, is expected to invade for any probability of recombination between loci \textbf{A} and \textbf{M}, $R$, when the net flow of recombinants is from the less fit (smaller $\lambda_{mi}$) to the more fit \textbf{A} background (making the terms inside the square brackets in Equations \ref{eq:neoYR} and \ref{eq:neoWR} negative). %\textcolor{red}{Q: is it definitely possible to have negative square brackets for a equilibria maintained by selection?}
%Assuming $g_A^* > 0$ and $g_a^* < 0$ without loss of generality, this is $-(1 - p^\female_X) w_{m,A} g_a^* < p^\female_X w_{m,a} g_A^*$.
%When the net flow of recombinants is from the more fit to the less fit haplotype, the new sex-determining allele can still invade when the rate of recombination between it and the selected locus is small enough. %\textcolor{red}{Q:Is it the case that sometimes the square brackets are positive and invasion occurs for $r_{AM}=1/2$? In which case it might be better to have slightly different phrasing here.}
%Equations \eqref{eq:neoYR} and \eqref{eq:neoWR} show that the new sex-determining allele, $m$, is expected to invade for any probability of recombination between loci \textbf{A} and \textbf{M}, $R$, when the net flow of recombinants is from the less fit (smaller $\lambda_{mi}$) to the more fit \textbf{A} background (making the terms inside the square brackets in Equations \ref{eq:neoYR} and \ref{eq:neoWR} negative).
%When the net flow of recombinants is from the more fit to the less fit haplotype, the new sex-determining allele can still invade when the rate of recombination between it and the selected locus is small enough. 
%To better understand when these scenarios are possible we next use knowledge of the equilibria and their stability under tight linkage and weak selection.

A rare neo-Y or neo-W will spread from a given ancestral equilibrium when the leading eigenvalue, $\lambda$, of the Jacobian matrix derived from the eight mutant recursion equations (given by \ref{eq:recursions}c,d,g,h), evaluated at the ancestral equilibrium, is greater than one.
However, because a neo-Y (neo-W) is always in males (females) and is epistatically dominant to the ancestral sex-determining locus, we need only two recursion equations (e.g., tracking the change in the frequency of neo-Y-$A$ and neo-Y-$a$ gametes from males) and thus the leading eigenvalue is the largest solution to a quadratic characteristic polynomial $\lambda^2 + b\lambda + c = 0$.
It can be shown (see supplementary Mathematica file) that the coefficients are $b= - (\lambda_{mA} + \lambda_{ma})+(\rho_{mA} + \rho_{ma})$ and $c = (\lambda_{mA}-\rho_{mA}) (\lambda_{ma}-\rho_{ma}) -\rho_{mA} \rho_{ma}$, where $\lambda_{mi}$ is the multiplicative growth rate of the frequency of mutants on background $i\in\{A,a\}$, without accounting for loss due to recombination, and $\rho_{mi}$ is the rate at which mutants on background $i\in\{A,a\}$ recombine onto the other \textbf{A} locus background in heterozygotes.
The leading eigenvalue is then greater than one whenever $\lambda_{mA}>1$ and $\lambda_{ma}>1$, less than one whenever $\lambda_{mA}<1$ and $\lambda_{ma}<1$, and greater than one whenever  $\lambda_{mA}>1$ or $\lambda_{ma}>1$ and $\rho_{ma} (\lambda_{mA}-1) + \rho_{mA} (\lambda_{ma}-1) > 0$.

For tight linkage between the ancestral sex-determining locus and the selected locus we can calculate each of these terms exactly, while for weak selection we take a Taylor series of the leading eigenvalue.

\subsubsection*{Recombination weak relative to selection (tight linkage between \textbf{A} and \textbf{X})}

Here, we explore the conditions under which a neo-W invades an XY system assuming that the \textbf{A} locus is initially in close linkage with the ancestral sex-determining region ($r \approx 0$). 
We disregard neo-Y mutations, which never spread given that the ancestral population is at a stable equilibrium.

Starting with the simpler equilibrium $(B)$, the terms of the characteristic polynomial are

\begin{subequations}\label{Binvasion}
\begin{align}
\lambda_{mA}&= (2 \alpha^\male)^{-1}\frac{w_{A}^\female \left[w_{A}^\male w_{AA}^\female \alpha^\male + 2 w_{a}^\male w_{Aa}^\female \alpha^\female(1-\alpha^\male)\right]}{ w_{A}^\male w_{A}^\female w_{AA}^\female } \\
\lambda_{ma}&= (2 \alpha^\male)^{-1} \frac{w_{a}^\female \left[ w_{a}^\male w_{aa}^\female (1-\alpha^\male) + 2 w_{A}^\male w_{Aa}^\female(1-\alpha^\female)\alpha^\male\right]}{w_{A}^\male w_{A}^\female w_{AA}^\female } \\
\rho_{mA}&= (2 \alpha^\male)^{-1}\frac{R w_{a}^\male w_{A}^\female w_{Aa}^\female \alpha^\female(1-\alpha^\male)}{w_{A}^\male w_{A}^\female w_{AA}^\female } \\
\rho_{ma}&= (2 \alpha^\male)^{-1}\frac{R w_{A}^\male w_{a}^\female w_{Aa}^\female (1-\alpha^\female)\alpha^\male}{w_{A}^\male w_{A}^\female w_{AA}^\female} 
\end{align}
\end{subequations}

%\noindent
%In this case, the zygotic sex ratio ($\zeta$) is given by the difference in haploid selection in males on $a$ (fixed on the Y) and $A$ (fixed on the X) alleles, i.e., there are more males than females if $\zeta = \alpha^\male w_A^\male / [(1-\alpha^\male)w_a^\male + \alpha^\male w_A^\male]<1/2$. %$\alpha^\male<1/2$ and/or $w_{a}^\male>w_{A}^\male$. 
%Populations with haploid selection for $a$ in males have a male biased zygotic sex ratio are thus more permissive to invasion by a neo-W ($\lambda_{mA}$ and $\lambda_{ma}$ larger).
%%In addition, the spread of neo-W alleles depends on the fitness of female diploids after mating with an X-$A$ or Y-$a$ male gamete, relative to the fitness of females in the population (all have genotype $AA$ and fitness $w_{AA}^\female$ at this equilibrium). 
%Haploid selection in males has a second effect; the spread rate of neo-W haplotypes is determined by their fitness in diploid females, which depends on their diploid genotype and thus on the male gamete they pair with. 
%Zygotes carrying dominant neo-W alleles will develop as females regardless of their genotype at the XY locus. 
%Therefore, neo-W females result from matings with either X-$A$ or Y-$a$ male gametes. 
%The relative proportion of these male gametes is determined by haploid selection in males; mating with a Y-$a$ male gamete is more likely if the $a$ allele is favoured during male gamete production or competition ($\zeta<1/2$). 
%Thus, neo-W females experience different diploid selection than XX females, and the extent of this difference depends on haploid selection in males.
%Furthermore, haploid selection in females can directly select upon neo-W-$A$ or neo-W-$a$ haplotypes. 
%A neo-W-$A$ female gamete has the same fitness during haploid competition as resident $A$-bearing female gametes. 
%On the other hand, neo-W-$a$ female gametes can be favoured or disfavoured during female haploid competition (favoured if $w_{a}^\female>w_{A}^\female$). 
%Meiotic drive in females ($\alpha^\female$) similarly affects the fitness of these neo-W haplotypes, except that it impacts both haplotypes as meiotic drive only occurs in heterozygotes and therefore does not occur in resident XX females (who are always homozygote $AA$).

Haploid selection impacts the spread of neo-W haplotypes in three ways.
Firstly, the zygotic sex ratio becomes male biased ($\zeta<1/2$; at equilibrium (B) the sex ratio is $\zeta = \alpha^\male w_A^\male / [(1-\alpha^\male)w_a^\male + \alpha^\male w_A^\male]$) when the $a$ allele (which is fixed on the Y) is favoured during competition among male gametes or by meiotic drive in males. 
This facilitates the spread of a neo-W because neo-W alleles cause the zygotes that carry them to develop as the rarer, female, sex. 
Secondly, haploid selection in males affects the diploid genotypes of females by altering the allele frequencies in the male gametes that female gametes pair with.
For instance, because an epistatically dominant neo-W always causes its carrier to become female, it creates females who carry either Y-$a$ or X genotypes from their father.
Thus, because when there is a polymorphism the X carries some non-zero frequency of $A$, haploid selection in males impacts the diploid genotypes of females (e.g., creating more $Aa$ females when drive in males favours Y-$a$).
How this affects the spread of the neo-W then depends on diploid and haploid selection in females.
Thirdly, female drive and gamete competition directly select on neo-W haplotypes.
Drive for $A$ in females favours neo-W-$A$ haplotypes, at a cost to neo-W-$a$ haplotypes, and vice-versa when there is drive for $a$.
The impact of this drive depends on how often XX and neo-W females are heterozygous.
Competition among female gametes acts similarly, and depends on the frequency of $A$ on resident X chromosomes (e.g., competition among eggs has no affect on the initial spread of the neo-W-$A$ haplotype when $A$ is fixed on the X).
Because haploid selection in females favours one neo-W haplotype at the expense of the other, recombination off the favoured background becomes more detrimental as it becomes more favoured. Thus higher rates of recombination between the neo-W and the selected locus, $R$, can lead to smaller leading eigenvalues when there is haploid selection in females.

The other terms in equations \eqref{Binvasion} are more easily interpreted if we assume that there is no haploid selection in either sex, in which case $\lambda_{mA}>1$ when $w_{Aa}^\female>w_{AA}^\female$ and $\lambda_{ma}>1$ when $(w_{Aa}^\female+w_{aa}^\female)/2>w_{AA}^\female$.
These conditions cannot be met under purely sexually-antagonistic selection, where $a$ is directionally favoured in males ($w_{AA}^\male>w_{Aa}^\male>w_{aa}^\male$) and $A$ is directionally favoured in females ($w_{AA}^\female>w_{Aa}^\female>w_{aa}^\female$).  
Essentially, the X is already as specialized as possible for the female beneficial allele (X$A$ is fixed), and the neo-W often makes daughters with the Y-a haplotype, increasing the flow of $a$ alleles into females, which reduces the fitness of those females.

If selection doesn't uniformly favour $A$ in females, however, neo-W-$A$ haplotypes and/or neo-W-$a$ haplotypes can spread ($\lambda_{mA}>1$ and/or $\lambda_{ma}>1$) at this equilibrium.
%If $\lambda_{mA}>1$ (requiring $w_{Aa}^\female>w_{AA}^\female$), the implication is that the neo-W can spread alongside the $A$ allele, despite the fact that the neo-W sometimes brings $Y-a$ haplotypes into females, because $a$ is favoured by selection in females despite $A$ being fixed on the X.
A neo-W can spread alongside the $A$ allele ($\lambda_{mA}>1$), despite the fact that a neo-W brings Y$a$ haplotypes into females, when $w_{Aa}^\female>w_{AA}^\female$.
In this case the $a$ allele is favoured by selection in females despite $A$ being fixed on the X.
For this equilibrium to be stable, X-$A$ must be sufficiently favoured in males to keep the frequency of X$A$ at one (specifically, from the stability conditions, we must have $w_{Aa}^\male/((w_{aa}^\male+w_{Aa}^\male)/2)>w_{Aa}^\female/w_{AA}^\female$).

Under this same condition, $w_{Aa}^\female>w_{AA}^\female$, the neo-W can also spread alongside the $a$ allele ($\lambda_{ma}>1$) if there is sufficiently strong underdominance in females ($w_{aa}^\female>w_{Aa}^\female$), such that $(w_{Aa}^\female+w_{aa}^\female)/2>w_{AA}^\female$.  
In this case, $a$ is not favored in females near the equilibrium where females are $AA$ (comparing $Aa$ to $AA$ genotypes) and yet the neo-W can spread with $a$ because it produces female $aa$ individuals by capturing Y-$a$ haplotypes.

When both haplotypes can spread on their own ($\lambda_{mA}>1$ and $\lambda_{ma}>1$), the neo-W invades regardless the recombination rate between it and the selected locus, $R$.
When neither haplotype can spread ($\lambda_{mA}<1$ and $\lambda_{ma}<1$) the neo-W can never invade.
And when only one haplotype can spread on its own the neo-W invades only when the rate of recombination onto the favourable background is sufficiently larger than the rate of recombination off this background (i.e., equation\ref{eq:lambdasGen} is satisfied).

%Haploid selection affects the conclusions outlined above by altering the zygotic sex ratio and exerting direct selection on neo-W haplotypes.
%The effect of sex-ratio selection is straightforward and can be seen in equation \eqref{Binvasion}, where male drive and gamete competition for $a$ (which is fixed on the Y and therefore causes a male bias) increases all four terms.
%Direct selection on neo-W haplotypes can also be seen in equation \eqref{Binvasion} by noticing that increasing female drive in favour of $A$ increases the neo-W-$A$ haplotype growth rate ($\lambda_{mA}$) while female drive and gamete competition in favour of $a$ increases the neo-W-$a$ haplotype growth rate ($\lambda_{ma}$).
%Gamete competition in females does not impact the spread of neo-W-$A$ haplotypes because they are on the same background as the resident X.
%Recombination between the selected locus and the neo-W complicates things slightly, as selective advantages for one neo-W haplotype during the haploid stage in females comes at a cost to the other, and thus recombination off the favoured background becomes more detrimental.

%\textcolor{red}{the versions for equil $(A)$ are not terrible but could be improved by cleverly re-writing with the $\phi$ and $\psi$ terms from above, I think...}

Similar equations can be derived for equilibrium (A) by subbing the equilibrium allele frequencies into Table \ref{tab:haplotype_growth}.

\begin{subequations}\label{Ainvasion}
\begin{align}
\lambda_{mA}&= \frac{a}{b} \left[w_{AA}^\female w_{Aa}^\male w_A^\male \alpha^\male \phi + 2 w_{Aa}^\female \alpha^\female w_a^\male \frac{c}{d}\right] / w_a^\female \\
\lambda_{ma}&= \frac{a}{b} \left[2 w_{Aa}^\female (1-\alpha^\female) w_{Aa}^\male w_A^\male \alpha^\male \phi+  w_{aa}^\female w_a^\male \frac{c}{d}\right] / w_A^\female \\
\rho_{mA}&= \frac{a}{b} R \left[ 2 w_{Aa}^\female \alpha^\female w_a^\male \frac{c}{d} \right] / w_a^\female\\ 
\rho_{ma}&= \frac{a}{b} R \left[ 2 w_{Aa}^\female (1-\alpha^\female) w_{Aa}^\male w_A^\male \alpha^\male \phi \right] / w_A^\female
\end{align}
\end{subequations}

\noindent
where

\begin{subequations}
\begin{align}
a &= w_a^\female \phi + w_A^\female \psi \\
b &= w_{AA}^\female \phi (2 w_{Aa}^\male w_A^\male \alpha_\male \phi) + w_{Aa}^\female \psi (2 w_{Aa}^\male w_A^\male \alpha_\male \phi + w_{AA}^\male w_a^\male \psi) + w_{aa}^\female \psi(w_{AA}^\male w_a^\male \psi) \\
c &= 2(w_{Aa}^\male\phi)^2(1-\alpha^\male)\alpha_\male + w_{Aa}^\male\phi(w_{AA}^\male \psi + w_{aa}^\male \psi \alpha^\male) + w_{aa}^\male \psi w_{AA}^\male \psi \\
d &= 2 w_{Aa}^\male \alpha^\male \phi + w_{aa}^\male \psi
\end{align}
\end{subequations}

As with equilibrium (B), haploid selection again modifies invasion fitnesses by altering the sex-ratio and the diploid genotypes of females and directly selecting upon female gametes.
The only difference is that resident XX females are no longer always homozygote $AA$ and males are no longer always heterozygote $Aa$.
Thus the effect of haploid selection in males is reduced, as is the difference in fitness between neo-W haplotypes and resident X haplotypes, as both can be on any diploid or haploid background.

The other terms are easier to interpret in the absence of haploid selection.
For instance, without haploid selection, the neo-W-$A$ haplotype spreads ($\lambda_{mA}>1$) if and only if

\begin{equation}\label{eq:BeqWAspread}
2(w_{Aa}^\female-w_{aa}^\female)w_{AA}^\male \psi^2 > (w_{AA}^\female-w_{Aa}^\female)w_{Aa}^\male \phi (\phi-\psi)
\end{equation}

\noindent  
where $\phi-\psi=w_{AA}^\female w_{Aa}^\male-w_{aa}^\female w_{aa}^\male$ and both $\phi$ and $\psi$ are positive when equilibrium (A) is stable. 
In contrast to equilibrium (B), a neo-W haplotype can spread under purely sexually-antagonistic selection  ($w_{AA}^\male>w_{Aa}^\male>w_{aa}^\male$ and $w_{AA}^\female>w_{Aa}^\female>w_{aa}^\female$).
In this case, the neo-W-$A$ haplotype can spread, despite producing a lot of $Aa$ daughters by obtaining the $a$ from Y-gametes, when $aa$ females, which the neo-W-$A$ never makes, are strongly selected against.
This can be intuited from the fact that \eqref{eq:BeqWAspread} will be more easily met when $w_{Aa}^\female-w_{aa}^\female\approx w_{Aa}^\female$ and $w_{AA}^\female-w_{Aa}^\female\approx 0$, implying $w_{aa}^\female \approx 0$ and $w_{Aa}^\female\approx w_{AA}^\female$ (although this is complicated by the fact that $w_{aa}^\female$ and $w_{Aa}^\female$ affect $\phi$ and $\psi$ too, the intuition holds).

Without haploid selection, the neo-W-$a$ haplotype spreads ($\lambda_{ma}>1$) if and only if

\begin{equation}\label{eq:BeqWaspread}
(w_{aa}^\female + w_{Aa}^\female-2w_{AA}^\female)w_{Aa}^\male \phi^2 + (w_{aa}^\female-w_{Aa}^\female)(w_{Aa}^\male+2w_{AA}^\male) \phi \psi >0
\end{equation}

\noindent
This condition cannot be met with purely sexually antagonistic selection (as both terms on the left-hand side would then be negative), but it can be met under other circumstances. 
For example, with overdominance in males there is selection for increased $A$ frequencies on X chromosomes in males, which are always paired with Y-$a$ haplotypes.
Then, directional selection for $a$ in females maintains a polymorphism at the $\textbf{A}$ locus on the X and by creating selection for decreased $A$ frequencies on X chromosomes in females.
This scenario selects for a modifier that increases recombination between the sex chromosomes \citep[e.g., blue region of Figure 2d in][]{Otto2014} and facilitates the spread of neo-W-$a$ haplotypes, which create more heterozygote and $aa$ females than ancestral X chromosomes do.

As with equilibrium (B), if both haplotypes can spread ($\lambda_{mA}>1$ and $\lambda_{ma}>1$) then the neo-W invades under any rate of recombination with the selected locus, $R\geq0$. 
In addition, even when only one haplotype can spread (e.g., under purely sexually-antagonistic selection $\lambda_{mA}>1$ and $\lambda_{ma}<1$), neo-W invasion can still occur under modest rates of recombination between the novel sex-determining and selected loci.

%Without haploid selection this reduces to
%\begin{subequations}
%\begin{align}
%\lambda_{mA}&= \frac{a}{b} \left[w_{AA}^\female w_{Aa}^\male \phi / 2 + w_{Aa}^\female \frac{c}{d}\right]  \\
%\lambda_{ma}&= \frac{a}{b} \left[w_{Aa}^\female w_{Aa}^\male \phi / 2 +  w_{aa}^\female \frac{c}{d}\right] \\
%\rho_{mA}&= \frac{a}{b} R \left[ w_{Aa}^\female \frac{c}{d} \right] \\ 
%\rho_{ma}&= \frac{a}{b} R \left[ w_{Aa}^\female w_{Aa}^\male \phi / 2 \right] 
%\end{align}
%\end{subequations}

%These simplify considerably when there is no haploid selection:
%
%\begin{subequations}
%\begin{align}
%\lambda_{mA}&= \frac{(\phi + \psi)[(w_{AA}^\female + w_{Aa}^\female ) w_{Aa}^\male \phi + 2 w_{Aa}^\female w_{AA}^\male \psi ]}{2 (w_{AA}^\female w_{Aa}^\male \phi^2 + \psi(w_{Aa}^\female(w_{AA}^\male + w_{Aa}^\male)\phi + w_{aa}^\female w_{AA}^\male \psi)}\\
%\lambda_{ma}&=  \frac{(\phi + \psi)[(w_{aa}^\female + w_{Aa}^\female ) w_{Aa}^\male \phi + 2 w_{aa}^\female w_{AA}^\male \psi ]}{2 (w_{AA}^\female w_{Aa}^\male \phi^2 + \psi(w_{Aa}^\female(w_{AA}^\male + w_{Aa}^\male)\phi + w_{aa}^\female w_{AA}^\male \psi)}\\
%\rho_{mA}&= R \frac{(\phi + \psi) w_{Aa}^\female w_{AA}^\male}{2 (w_{AA}^\female w_{Aa}^\male \phi^2 + \psi(w_{Aa}^\female(w_{AA}^\male + w_{Aa}^\male)\phi + w_{aa}^\female w_{AA}^\male \psi)} \\ 
%\rho_{ma}&= R \frac{(\phi + \psi) w_{Aa}^\female w_{Aa}^\male}{2 (w_{AA}^\female w_{Aa}^\male \phi^2 + \psi(w_{Aa}^\female(w_{AA}^\male + w_{Aa}^\male)\phi + w_{aa}^\female w_{AA}^\male \psi)}\\
%\end{align}
%\end{subequations}
%
%\begin{subequations}
%\begin{align}
%\lambda_{mA}&= [( w_{AA}^\female + w_{Aa}^\female ) w_{Aa}^\male \phi + 2 w_{Aa}^\female w_{AA}^\male \psi] a \\
%\lambda_{ma}&=  [(w_{aa}^\female + w_{Aa}^\female ) w_{Aa}^\male \phi + 2 w_{aa}^\female w_{AA}^\male \psi ] a\\
%\rho_{mA}&= R w_{Aa}^\female (w_{Aa}^\male \phi + 2 w_{AA}^\male \psi) c \\ 
%\rho_{ma}&= R w_{Aa}^\female w_{Aa}^\male \phi c \\
%\end{align}
%\end{subequations}
%\noindent
%where 
%$c = \frac{\phi + \psi}{2 [w_{AA}^\female w_{Aa}^\male \phi^2 + \psi(w_{Aa}^\female(w_{AA}^\male + w_{Aa}^\male)\phi + w_{aa}^\female w_{AA}^\male \psi]}$ and $\phi$ and $\psi$ are evaluated with $\alpha^{\Hermaphrodite}=1/2$ and $w_a^\Hermaphrodite=w_A^\Hermaphrodite=1$.

%With $R=0$ the two eigenvalues are $\lambda_{mA}$ and $\lambda_{ma}$.
%The neo-W can then invade on the $A$ background when $\lambda_{mA}>1$ and the neo-W can invade on the $a$ background when $\lambda_{ma}>1$.
%Even in this case it is difficult to see from equation \ref{Ainvasion} when the neo-W will invade.
%Below we discuss why $\lambda_{mA}$ and $\lambda_{ma}$ might be larger than one and confirm the logic numerically.
%
%
%For simplicity, consider the case of no haploid selection.
%A polymorphism at the $\mathbf{A}$ locus then requires sexually-antagonistic selection \citep{vanDoorn:2007eu}.
%Given that $a$ is fixed on the Y, one can consider the $a$ allele `male beneficial' and the $A$ allele to be `female beneficial'. 
%Because X chromosomes spend some time in males the frequency of $A$ on the X can reach an intermediate equilibrium (equilibrium A).
%
%We can therefore reason that when a neo-W invades equilibrium (A), it is likely to spread on the $A$ background as W chromosomes are always in females.
%The neo-W can outcompete the ancestral, `comprimising' X because it can specialize on females by fixing the `female beneficial' allele.  
%We call this the `sex-specialist hypothesis'.
%This hypothesis should only work when the benefit of specializing outweighs the cost of bringing Y chromosomes into females (the neo-W is epistatically dominant over the X/Y locus).
%
%One could also reason that, when the neo-W spreads, it does so precisely because it brings Y chromosomes, and their associated alleles, into females, something the X cannot do.
%However, this requires a somewhat particular (perhaps `perverse') equilibrium, where XX females do worse than XY females despite associations on the X and Y being built up by selection.
%One example of such a perverse equilibrium is produced by overdominance in both sexes.
%Overdominance in females selects for an allele frequency of 0.5 on the X ($A$ and $a$ equally frequent, maximizing the number of heterozygotes). 
%However, given that $a$ is fixed on the Y, overdominance in males selects for the X to become more associated with the $A$. 
%Thus, a compromising equilibrium is reached on the X chromosome and females fitness suffers due to the overproduction of $AA$ homozygotes.
%This sets up a scenario in which a neo-W can invade on the $a$ background, despite the fact that $a$ is fixed on the Y and in that sense `male beneficial'.
%We call this the `perverse hypothesis'.
%
%Figure \ref{fig:Region_Plots} shows where the sex-specialist (red) and perverse (blue) hypotheses are valid.
%The left panel demonstrates that both hypotheses can be simultaneously valid with overdominance in both sexes and sex differences in selection, which overlaps the parameter region where there is selection for increased recombination between the sex chromosomes \citep[compare with figure 2a in][]{Otto2014}.
%The right panel shows that the perverse hypothesis is no longer valid when selection is exactly the same in both sexes, despite the fact that there is selection for increased recombination between the sex chromosomes with sufficiently strong overdominance \citep[compare with figure 3 in][]{Otto2014}.
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%Sex-specialist vs perverse hypotheses
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{figure}[!h]
%\centering
%\includegraphics[width=0.45\linewidth]{regionplot_2a}
%\includegraphics[width=0.45\linewidth]{regionplot_3}\\
%\caption{
%Numerical exploration of where sex-specialist (red) and perverse (blue) hypotheses are valid, i.e., where an internally stable resident equilibrium can be invaded by a neo-W that is perfectly linked with the allele not fixed on the Y (red) or the allele that is fixed on the Y (blue). 
%\textbf{(Left)}
%The perverse equilibrium is only valid with overdominance while the sex-specialist hypothesis is valid over the entire range shown.
%Both hypotheses are simultaneously valid over much of the parameter space where increased recombination between the sex chromosomes is selected \citep[][figure 2a]{Otto2014}.
%\textbf{(Right)}
%Only the sex-specialist hypothesis is valid with no sex-differences in selection, despite selection for increased recombination between the sex-chromosomes with strong overdominance \citep[][figure 3]{Otto2014}.
%Parameters: $w_{Aa}^\Hermaphrodite = 1$, $w_a^\Hermaphrodite = w_A^\Hermaphrodite = 1$, $\alpha^\Hermaphrodite = 1/2$, $r=R=0$, and in left panel $w_{aa}^\female = w_{AA}^\female = 0.6$.
%}
%\label{fig:Region_Plots}
%\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection*{Selection weak relative to recombination (weak selection)}

With weak selection the leading eigenvalue, $\lambda$, for any $k$, is given up to order $\epsilon^2$ by equation \ref{eq:lambda_ESD_k}.
Scenarios leading to $\lambda>1$ are discussed in the main text.

\newpage

%\section*{Supplementary Figures}

\end{document}