### 2018-02-04 cover.tex
cover and route 1

 \documentclass[12pt,letterpaper]{article}
\usepackage{fontenc}
\usepackage[english]{babel}

\usepackage{fullpage}

\usepackage[pdftex]{graphicx}

\begin{document}

\thispagestyle{empty}

\noindent Michael F.\ Scott\\
Genetics, Evolution \& Environment\\
University College London Genetics Institute\\
2nd Floor, Darwin Bldg., Gower St., London, WC1E 6BT, UK\\
 m.f.scott@ucl.ac.uk\\
\smash{\hspace{12cm}\includegraphics[scale=0.5]{UCL.png}}

%\noindent Matthew M.\ Osmond\\
%Department of Zoology\\
%University of British Columbia\\
%6270 University Blvd. \\
%Vancouver, BC V6T 1Z4, Canada\\
%mmosmond@zoology.ubc.ca\\
%\smash{\hspace{11.5cm}\includegraphics[scale=0.25]{IMAGES/UBC.png}}

\noindent Dear Editor, \hfill \today \\

Please find the attached manuscript, entitled ``Haploid selection, sex ratio bias, and transitions between sex-determination systems'', which we submit to you for consideration for publication in \textit{PLoS Biology}.
\\

%findings;why important
[why this is cool]
\\

%Based upon their expertise in theoretical quantitative genetics in changing environments, we suggest Dr.\ Luis-Miguel Chevin$^1$, Dr.\ Richard Gomulkiewicz$^2$, and Dr.\ Reinhard B\"urger$^3$ as potential reviewers.

%This work is original and is not under consideration for publication elsewhere. 
%This work is original, unpublished and is not under consideration for publication elsewhere. %H
%All authors have agreed to the submission of the paper to \textit{Evolution}. %H

We hope you agree that this work is of broad interest to the readership of \textit{PLoS Biology} and we eagerly anticipate your response.\\

Sincerely,

[signature]\\
%\includegraphics[height=1.5cm]{IMAGES/OsmondSignature.pdf}
\indent Michael Scott, Matthew Osmond, Sarah Otto

%reviewer emails
%\small{$^1$ luis-miguel.chevin@cefe.cnrs.fr;  $^2$ gomulki@wsu.edu; $^3$ reinhard.buerger@univie.ac.at}

\end{document}